---
title: "Introduction to Inference: Confidence Intervals"
format: 
  live-html:
    webr:
      packages: 
        - ggformula
        - mosaic
        - tidyverse
      repos:
        - https://projectmosaic.r-universe.dev
      resources:
        - https://raw.githubusercontent.com/jcanner/math-320/refs/heads/master/note-outlines/8-intro-to-stats/king.csv
      cell-options:
        warning: true
        completion: false
execute: 
  warning: false 
  message: false
  echo: true
embed-resources: false
code-copy: true
code-overflow: wrap
toc: true
number-sections: true
css: ../styles.css
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}
{{< include ../_extensions/r-wasm/live/_gradethis.qmd >}}

```{webr}
#| label: theme
#| include: false
#| edit: false
library(ggplot2)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

```{r}
#| label: setup-checkdown
#| include: false
library(checkdown)
library(ggformula)
library(mosaic)
library(tidyverse)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

---
title: "Introduction to Statistical Hypothesis Testing"
author: "Steven B. Kim"
format: 
  html:
    css: styles.css
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    toc: true
    toc-location: left
    number-sections: true
    html-math-method: katex
mainfont: serif
editor: visual
---

In statistics, a **hypothesis** refers to an assumption about the population or an unknown parameter (any quantity which describes the population). For example, let $\theta$ denote the proportion of boys among newborn babies in the United States. Then $\theta = 0.5$ is a hypothesis stating that the population proportion is 0.5 (i.e., the probability that a mother gives birth to a boy is 0.5). Another competing hypothesis is $\theta > 0.5$. This hypothesis states that the probability that a mother gives birth to a boy is greater than 0.5.

## Two Kinds of Hypothesis {#sec-hypothesis}

In statistical hypothesis testing, we specify two competing hypotheses: **null hypothesis** and **alternative hypothesis**. The null hypothesis is denoted by $H_0$, and the alternative hypothesis is denoted by $H_1$. The procedure of hypothesis testing is developed based on the assumption of $H_0$. If data suggest strong evidence against $H_0$ in favor of $H_1$, we conclude $H_1$. Otherwise, we conclude $H_0$.

**Note:** Conclusion of $H_1$ does not imply that $H_1$ is true. It implies that observed data (a sample) suggest strong evidence against $H_0$ in favor of $H_1$. Similarly, conclusion of $H_0$ does not imply that $H_0$ is true. It implies that observed data is not strong enough to reject the assumption of $H_0$.

## Analogy: Presumption of Innocence in U.S. Criminal Trials {#sec-analogy}

Statistical hypothesis testing described in @sec-hypothesis is somewhat similar to criminal trials in the United States. There are two competing hypotheses in a criminal trial. The first hypothesis is that the defendant is innocent, and it is assumed to be true until proven guilty (**presumption of innocence**). The second hypothesis is that the defendant is guilty. The prosecutor must provide strong evidence against the presumption of innocence in order to reject the presumption and conclude guilty. According to our criminal justice system, after seeing all evidence provided by the prosecutor, all 12 jurors must agree that the defendant is guilty in order to conclude guilty. Otherwise, we conclude that the defendant is innocent. (It is not because we truly believe that the defendant is innocent. It is because the evidence is not strong enough to reject the presumption of innocence.)

## Three Kinds of Alternative Hypothesis

Let $\theta$ be a parameter to be tested. The null hypothesis is simply stated as $H_0$: $\theta = \theta_0$ where $\theta_0$ is referred to as the **null value**. In the previous example, we may set $H_0$: $\theta = 0.5$ (i.e., the probability of giving birth to a boy is 0.5). There are three forms of alternative hypothesis $H_1$

- $H_1$: $\theta > \theta_0$ which is referred to as the **right-sided test**.

- $H_1$: $\theta < \theta_0$ which is referred to as the **left-sided test**.

- $H_1$: $\theta \ne \theta_0$ which is referred to as the **two-sided test**.

Given $H_0$: $\theta = \theta_0$, the direction of $H_1$ depends on the researcher's question of interest. In the previous example, we may set $H_0$: $\theta = 0.5$ and $H_1$: $\theta > 0.5$ (right-sided test) because the sex ratio at birth is known to be about 105 to 100 (boys to girls). The two competing hypotheses $H_0$ and $H_1$ are assumptions, and we need data (a sample) to conclude $H_0$ or $H_1$.

## Test Statistic {#sec-teststat}

A statistic (numeric summary of data) is called a **test statistic** if it is used to conduct statistical hypothesis testing. In order to test $H_0$: $\theta = 0.5$ and $H_1$: $\theta > 0.5$, we need to analyze a categorical variable with two levels: boy and girl (male and female) at birth. For the purpose of demonstration, let us revisit the $\texttt{king.csv}$ dataset, a simple random sample of 2,500 mothers who gave births in 2001 in King County. Below are descriptive statistics and data visualization relevant to our hypothesis testing.

```{r, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width = 5, out.height = "50%", out.width = "50%", fig.alt = "Barplot by proportion"}
library(mosaic)
king = read.csv("Data/king.csv") ### read csv file
dim(king) ### dimension of data
head(king, 5) ### view first 5 lines
count = tally(~gender, data = king, format = "count")
prop = tally(~gender, data = king, format = "proportion")
count
prop
barplot( prop, 
         beside = TRUE,
         col = c("pink", "blue"),
         main = "Newborn Babies in 2001 in King County",
         ylab = "Proportion", 
         ylim = c(0,1),
         names = c("Girls (1209)", "Boys (1291)") )
```

In the sample, we observed 1209 girls ($\texttt{F}$) and 1291 boys ($\texttt{M}$) out of $n = 2500$ newborn babies. The unknown parameter $\theta$, the probability of giving birth to a boy, can be estimated by the sample proportion of boys, $\bar{X} = \frac{1291}{2500} = 0.5164$, and we can calculate a test statistic as follows:

$$
Z = \frac{\bar{X} - \theta_0}{ \sqrt{ \frac{\theta_0 (1 - \theta_0)}{n} } } = \frac{0.5164 - 0.5}{ \sqrt{ \frac{0.5(1 - 0.5)}{2500} } } = 1.64 \, .
$$

```{r}
n = dim(king)[1] ### the number of rows in the data (sample size)
xbar = 1291 / n ### sample proportion of boys
theta0 = 0.5 ### null value
se = sqrt( theta0 * (1 - theta0) / n ) ### standard error
z = (xbar - theta0) / se ### z statistic
z

```

This test statistic is sometimes referred to as a **Z statistic**. The numerator of $Z$ measures the discrepancy between the sample proportion ($\bar{X} = 0.5164$) and the null hypothesis ($\theta_0 = 0.5$). The denominator $\texttt{SE} = \sqrt{ \frac{\theta_0 (1 - \theta_0) }{n} }$ is called the **standard error** of $\bar{X}$. The standard error measures uncertainty in the sample proportion given the null hypothesis ($\theta_0 = 0.5$) and sample size ($n = 2500$). In our case, a large positive value of $Z$ is regarded as strong evidence against $H_0$: $\theta = 0.5$ in favor of $H_1$: $\theta > 0.5$, and it can be attained when $\bar{X} > \theta_0$ and $\texttt{SE}$ is small. A natural question is how large $Z$ should be in order to reject $H_0$ and conclude $H_1$. This question will be addressed in @sec-rule.

## Two Kinds of Mistake in Statistical Hypothesis Testing

There are two kinds of mistake in statistical hypothesis testing. The first kind of mistake is concluding $H_1$ when $H_0$ is true, and it is referred to as the **Type I error**. The second kind of mistake is concluding $H_0$ when $H_1$ is true, and it is referred to as the **Type II error**. Since the process of hypothesis testing is built upon the assumption of $H_0$, we conclude $H_0$ or $H_1$ by fixing the probability of committing Type I error. This fixed probability is called the **significance level**, and denoted by $\alpha$. A typical choice is $\alpha = 0.01$, $\alpha = 0.05$, or $\alpha = 0.1$. If we fix $\alpha = 0.05$, the underlying interpretation is that we allow a 5% chance of committing Type I error (concluding $H_1$ when $H_0$ is true).

## Rejection Rule {#sec-rule}

Under the assumption of $H_0$, the reference probability model for the Z statistic is the standard normal distribution, and we write $Z \sim N(0,1)$. It is also called the **null distribution** of $Z$ (the sampling distribution of $Z$ under the assumption of $H_0$).

```{r, echo = FALSE, fig.height = 4.5, fig.width = 5, out.height = "50%", out.width = "50%", fig.alt = "Null distribution of Z statistic"}
mu = 0
sigma = 1
x = seq( mu - 4 * sigma, mu + 4 * sigma, 0.01 )
y = dnorm( x, mu, sigma )
plot( x, y, type = "l", col = 1, lty = 1, 
      main = "Null Distribution of Z", 
      xlab = "Value of Z", 
      ylab = "",
      xlim = c( mu - 3.5 * sigma, mu + 3.5 * sigma ), 
      ylim = c( 0, 0.4 ), 
      axes = FALSE )
axis( 1, seq( mu - 5 * sigma, mu + 5 * sigma, sigma ) )
abline( h = 0, col = 8 )
```

After fixing a significance level $\alpha$, we conclude $H_1$ when:

- [Right-sided test]{.underline}: $Z \ge c$ where $c$ is calculated by $\texttt{qnorm}(1 - \alpha)$.

- [Left-sided test]{.underline}: $Z \le c$ where $c$ is calculated by $\texttt{qnorm}(\alpha)$.

- [Two-sided test]{.underline}: $\lvert Z \rvert \ge c$ where $c$ is calculated by $\texttt{qnorm}(1 - \alpha/2)$.

Otherwise, we conclude $H_0$. The value of $c$ is called a **critical value**, and it is like a threshold to determine $H_1$ or $H_0$ at a fixed significance level $\alpha$.

Remember that the process of hypothesis testing is formulated under the assumption of $H_0$. Therefore, when we conclude $H_0$, we carefully state that it is insufficient evidence to reject $H_0$ at a significance level $\alpha$. We shall [not]{.underline} state that it is sufficient evidence to conclude $H_0$. (Recall the analogy in Section @sec-analogy. In a criminal trial, when we conclude innocence, it is because of insufficient evidence to reject the presumption of innocence. It does not imply that the prosecutor provided sufficient evidence for innocence.)

In @sec-teststat, supposed we fixed a significance level of $\alpha = 0.05$, and we calculated $Z = 1.64$. We can calculate the critical value $c$ as follows.

```{r}
alpha = 0.05 ### significance level
c = qnorm(1 - alpha) ### critical value
c
```

It was close, but $Z = 1.64$ is less than $c = 1.6449$. Therefore, we conclude $H_0$: $\theta = 0.5$. In words, the sample of 2,500 newborn babies is not sufficient evidence to conclude that the probability of giving birth to a boy is greater than 0.5. The parameter $\theta$ is estimated as $\bar{X} = 0.5164$, and we would need a larger sample size to conclude $H_1$: $\theta > 0.5$ if it is true.

## P-value

In statistics, the **p-value** is broadly defined as the probability of observing an observed outcome or a more unlikely outcome in favor of $H_1$ under the assumption of $H_0$. After calculating $Z = \texttt{z}$, the calculation of p-value depends on the direction of $H_1$ as follows:

- [Right-sided test]{.underline}: p = $1 - \texttt{pnorm}(\texttt{z})$

- Left-sided test: p = $\texttt{pnorm}(\texttt{z})$

- Two-sided test: p = $2 * \texttt{pnorm}(\texttt{z})$ if $\texttt{z} < 0$ or p = $2 * (1 - \texttt{pnorm}(\texttt{z}))$ if $\texttt{z} > 0$

```{r, echo = FALSE, fig.height = 3, fig.width = 10, out.height = "100%", out.width = "100%", fig.alt = "P-value calculation"}
par( mfrow=c( 1, 3 ) )

###

mu = 0 ### model mean
sigma = 1 ### model standard deviation
x = seq( -4, 4, 0.001 )

plot( x, dnorm( x, mu, sigma ), type = "l", axes = FALSE, xlim = c( -4, 4 ), xlab = "Z Statistic", ylab = "", main = "Left-sided Test" )
abline( h = 0 )

a = -5; b = -1.5
grid = seq( a, b, 0.01 )
for ( i in 1:length(grid) ) {
   x.temp = c( grid[i], grid[i] )
   y.temp = c( 0, dnorm( grid[i], mu, sigma ) )
   lines( x.temp, y.temp, col = "gold" ) }

z = b
axis( 1, c( -10, 0, z, 10 ), label = c( "", "0", "z", "" ) )

legend( "topright", lwd = 5, col = "gold", legend = "P-value", bty = "n" )

###

mu = 0 ### model mean
sigma = 1 ### model standard deviation
x = seq( -4, 4, 0.001 )

plot( x, dnorm( x, mu, sigma ), type = "l", axes = FALSE, xlim = c( -4, 4 ), xlab = "Z Statistic", ylab = "", main = "Right-sided Test" )
abline( h = 0 )

a = 1.5; b = 5
grid = seq( a, b, 0.01 )
for ( i in 1:length(grid) ) {
   x.temp = c( grid[i], grid[i] )
   y.temp = c( 0, dnorm( grid[i], mu, sigma ) )
   lines( x.temp, y.temp, col = "gold" ) }

z = a
axis( 1, c( -10, 0, z, 10 ), label = c( "", "0", "z", "" ) )

legend( "topright", lwd = 5, col = "gold", legend = "P-value", bty = "n" )

###

mu = 0 ### model mean
sigma = 1 ### model standard deviation
x = seq( -4, 4, 0.001 )

plot( x, dnorm( x, mu, sigma ), type = "l", axes = FALSE, xlim = c( -4, 4 ), xlab = "Z Statistic", ylab = "", main = "Two-sided Test" )
abline( h = 0 )

a = 1.5; b = 5
grid = seq( a, b, 0.01 )
for ( i in 1:length(grid) ) {
   x.temp = c( grid[i], grid[i] )
   y.temp = c( 0, dnorm( grid[i], mu, sigma ) )
   lines( x.temp, y.temp, col = "gold" ) }

z = a
axis( 1, c( -10, -z, 0, z, 10 ), label = c( "", "-|z|", "0", "|z|", "" ) )

a = -5; b = -1.5
grid = seq( a, b, 0.01 )
for ( i in 1:length(grid) ) {
   x.temp = c( grid[i], grid[i] )
   y.temp = c( 0, dnorm( grid[i], mu, sigma ) )
   lines( x.temp, y.temp, col = "gold" ) }

legend( "topright", lwd = 5, col = "gold", legend = "P-value", bty = "n" )
```

Instead of comparing the value of $Z$ to the critical value $c$ as in @sec-rule, we can compare p-value to the significance level $\alpha$. If the value of $p$ is less than or equal to $\alpha$, we conclude $H_1$. Otherwise, we conclude $H_0$.

In the previous example, we can calculate the p-value for the right-sided test as follows.

```{r}
n = dim(king)[1] ### the number of rows in the data (sample size)
xbar = 1291 / n ### sample proportion of boys
theta0 = 0.5 ### null value
se = sqrt( theta0 * (1 - theta0) / n ) ### standard error
z = (xbar - theta0) / se ### z statistic
z
p = 1 - pnorm(z) ### p-value for right-sided test
p
```

It was close, but p = 0.0505 is greater than $\alpha = 0.05$. Therefore, we conclude $H_0$. We interpret p = 0.0505 as follows. If we repeat simple random sampling of size $n = 2500$, the probability of observing $Z = 1.64$ or $Z > 1.64$ is p = 0.0505 if $H_0$: $\theta = 0.5$ is true. This result is not statistically significant enough at the fixed significance level of $\alpha = 0.05$, so we conclude $H_0$ at $\alpha = 0.05$.

## General Procedure of Statistical Hypothesis Testing

In summary, the procedure of statistical hypothesis testing is as follows:

1.  Choose a parameter of interest, and state $H_0$ and $H_1$ in terms of the parameter. It depends on the research question.
2.  Fix a significance level $\alpha$. It is a small fixed probability of committing Type I error (concluding $H_1$ when $H_0$ is true).
3.  Calculate a test statistic and p-value under the assumption of $H_0$.
4.  Conclude $H_0$ or $H_1$. You may compare the calculated test statistic to $c$ or the p-value to $\alpha$.
5.  State the conclusion in the context.

**Remark:** In this chapter, we discussed the procedure when a parameter of interest is $\theta$, the unknown probability of an event (the population proportion of a certain group). The procedure is known to be valid when (1) we have a simple random sample of size $n$ and (2) the sample size $n$ is as large as follows: $n \theta_0 \ge 10$ or $n(1 - \theta_0) \ge 10$. In our example, $n = 2500$ and $\theta_0 = 0.5$, so $n$ is large enough ($n\theta_0 = 1250 \ge 10$ and $n(1-\theta_0) = 1250 \ge 10$).

## Warm-up

*You must input your answers in the [Google form](https://docs.google.com/forms/d/e/1FAIpQLSddNCcu36e40Xo51e-cbJvQk_h5Us7rr_jeS4CVq52UQMxUZA/viewform?usp=dialog). All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Any typo will result in point deduction in the auto-grading system. Any submission after the due date specified in Canvas will not be graded.*

**Problem 1.** Fill in the blank.

a\. A \[h...\] is an assumption about the population or an unknown parameter.

b\. A \[t... s...\] is a statistic which is used for statistical hypothesis testing.

c\. A \[n... h...\] is assumed to be true in the procedure of statistical hypothesis testing.

d\. If data is strongly against the null hypothesis in favor of the \[a... h...\], we conclude the \[a... h...\].

e\. \[...\] refers to conclusion of $H_1$ when $H_0$ is true.

f\. \[...\] refers to conclusion of $H_0$ when $H_1$ is true.

g\. A \[s... l...\] is denoted by $\alpha$, and it is a small fixed probability of committing the Type I error.

h\. We compare a test statistic to the \[c... v...\] to conclude $H_0$ or $H_1$.

i\. We can a p-value to the \[...\] to conclude $H_0$ or $H_1$.

**Problem 2.** True or false?

a\. If $H_0$ is true, it is possible to conclude $H_1$.

b\. If $H_1$ is true, it is possible to conclude $H_0$.

c\. If we fix $\alpha = 0.05$, it implies that we allow a probability of 0.05 for concluding $H_0$ when $H_1$ is true.

d\. When a p-value is less than $\alpha$, we conclude $H_1$.

e\. If $H_0$ is concluded, it implies that $H_0$ is true.

f\. If $H_1$ is concluded, it implies that $H_1$ is true.

g\. If a p-value is 0.09, it implies that the probability that $H_0$ is true is 0.09.

**Problem 3.** Suppose a researcher calculated a Z statistic of 2.1. Calculate the p-value for:

a\. right-sided test

b\. left-sided test

c\. two-sided test

**Problem 4.** Multiple-choice: Suppose we test $H_0$: $\theta = 0.7$ versus $H_1$: $\theta > 0.7$ (right-sided test). Which of the following provides the strongest evidence for $H_1$?

\(A\) $\bar{X} = 0.75$ and $n = 100$

\(B\) $\bar{X} = 0.72$ and $n = 1000$

\(C\) $\bar{X} = 0.71$ and $n = 10000$

\(D\) $\bar{X} = 0.7$ and $n = 500$

\(E\) $\bar{X} = 0.8$ and $n = 50$

**Problem 5.** Multiple-choice: Which case results in the smallest p-value in Problem 4?

**Problem 6.** Multiple-choice: Suppose we test $H_0$: $\theta = 0.3$ versus $H_1$: $\theta < 0.3$ (left-sided test). Which of the following results in the smallest p-value?

\(A\) $\bar{X} = 0.35$ and $n = 500$

\(B\) $\bar{X} = 0.3$ and $n = 100$

\(C\) $\bar{X} = 0.29$ and $n = 100$

\(D\) $\bar{X} = 0.25$ and $n = 100$

\(E\) $\bar{X} = 0.25$ and $n = 200$

**Problem 7.** Multiple-choice: If we test $H_0$: $\theta = 0.3$ versus $H_1$: $\theta \ne 0.3$ (two-sided test), which case in Problem 6 results in the smallest p-value?

**Problem 8.** Multiple-choice: Which of the following can be a negative value?

\(A\) sample proportion

\(B\) p-value

\(C\) Z statistic

\(D\) standard error

\(E\) There are more than one answer among (A) to (D)

## Assignment

***Instruction 1:** You must input your answers in the [Google form](https://docs.google.com/forms/d/e/1FAIpQLSfT0OoWDf6dfBKlbsZh3GO6inPMKlSXWR2Y1oCdOcMqxrN27w/viewform?usp=dialog). All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Any typo will result in point deduction in the auto-grading system. Any submission after the due date specified in Canvas will not be graded.*

***Instruction 2:** You must submit an html file generated by Quarto in Canvas. Regardless of your answers in the Google form, 25% of the assignment will be graded based on the quality of the html document (e.g., completeness, organization, easiness to read, use of complete sentences and accurate notation). If an R code which is not demonstrated in the course material is used, it will be regarded as use of unacceptable sources and/or aids, and it will result in zero in the assignment. Such a case may be reported according to the [CSUMB Academic Integrity Policy](https://csumb.edu/policy/academic-integrity-policy/).*

**Problem 1.** Let $\theta$ be the U.S. President's approval rating (i.e., the population proportion of Americans who think the President is doing a good job). Suppose a 50% approval rating is a benchmark that the President is doing a good job, and it has been believed that the approval rating is below 50% since the election. In May 2025, Gallup randomly sampled 1,003 adults, and 431 of them responded that the President is doing a good job.

a\. What is the null value in this context? (Numeric answer)

b\. Multiple-choice: Given the above description, what kind of statistical hypothesis testing do we want to do?

\(A\) right-sided test

\(B\) left-sided test

\(C\) two-sided test

c\. Calculate the sample proportion of people who approved.

d\. Calculate the standard error.

e\. Calculate the Z statistic.

f\. Calculate the p-value.

g\. Multiple-choice: At a significance level of $\alpha = 0.05$, what do you conclude?

\(A\) $H_0$

\(B\) $H_1$

\(C\) Inconclusive

h\. Multiple-choice: The significance level $\alpha = 0.05$ implies that we allowed a probability of 0.05 for...

\(A\) concluding $\theta = 0.5$ when $\theta > 0.5$ is true.

\(B\) concluding $\theta = 0.5$ when $\theta < 0.5$ is true.

\(C\) concluding $\theta = 0.5$ when $\theta \ne 0.5$ is true.

\(D\) concluding $\theta > 0.5$ when $\theta = 0.5$ is true.

\(E\) concluding $\theta < 0.5$ when $\theta = 0.5$ is true.

\(F\) concluding $\theta \ne 0.5$ when $\theta = 0.5$ is true.

i\. Multiple-choice: Which of the following statement is the most appropriate concluding statement?

\(A\) The approval rating is 50%.

\(B\) The approval rating is below 50%.

\(C\) The sample provides insufficient evidence to conclude that the approval rating is 50%.

\(D\) The sample provides sufficient evidence to conclude that the approval rating is 50%.

\(E\) The sample provides insufficient evidence to conclude that the approval rating is below 50%.

\(F\) The sample provides sufficient evidence to conclude that the approval rating is below 50%.

**Problem 2.** A child, who loves outdoor activities on weekend, believes that it rains more often on weekend (Saturdays and Sundays) than on weekdays (Monday to Friday). Of the last 100 rainy days, 33 days were on Saturdays and Sundays. Does this weather record support the child's hypothesis? Let $\theta$ be the true proportion of Saturdays and Sundays among all rainy days.

a\. What is the null value in this context? (Numeric answer)

b\. Multiple-choice: Which of the following represents the alternative hypothesis?

\(A\) right-sided test

\(B\) left-sided test

\(C\) two-sided test

c\. Calculate the sample proportion?

d\. Calculate the standard error.

e\. Calculate the Z statistic.

f\. Calculate the p-value.

g\. Multiple-choice: At a significance level of $\alpha = 0.1$, what do you conclude?

\(A\) $H_0$

\(B\) $H_1$

\(C\) Inconclusive

h\. Multiple-choice: In this hypothesis testing, the null hypothesis $H_0$ is that the chance of rain is the same everyday, and the alternative hypothesis $H_1$ is that the chance of rain is higher on Saturday and Sunday than on Monday to Friday. If the p-value is correctly calculated in part f, the p-value is greater than $\alpha = 0.1$. Which of the following statement is the most appropriate concluding statement?

\(A\) Assuming $H_0$ is true, the data provide insignificant evidence for $H_0$.

\(B\) Assuming $H_0$ is true, the data provide insignificant evidence against $H_0$.

\(C\) Assuming $H_1$ is true, the data provide insignificant evidence for $H_1$.

\(D\) Assuming $H_1$ is true, the data provide insignificant evidence against $H_1$.

**\*Problem 3.** Suicidal behavior is a leading cause of death and injury in the world (Klonsky, 2016). While women are more likely to experience suicidal ideation and attempt, men are more likely to die per attempt. It is known as the [Gender Paradox of Suicide](https://cams-care.com/resources/educational-content/the-gender-paradox-of-suicide/). According to the Centers for Disease Control and Prevention (CDC), men were four times as likely as women to commit suicide in 2021. Download $\texttt{ShandongChina.csv}$. It is a record of suicide attempts in 2009, 2010, and 2021 in a rural population in Shandong, China (Sun et al., 2015).

Our goal is to answer two questions: (1) Are women more likely to attempt suicide than men? (2) Are there more men who died as a result of suicide attempt than women? Assume that 50% are men and 50% are women in the general population of Shandong, China. You may modify the following R code for answering the following questions.

```{r, warning = FALSE, message = FALSE}
library(mosaic)
shandong = read.csv("Data/ShandongChina.csv") ### data for question (1)
dim(shandong) ### dimension of data
head(shandong, 5) ### view first 5 lines

shandong2 = subset(shandong, Died == "yes") ### subset the data (those who died) for question (2)
dim(shandong2)
head(shandong2, 5)
```

*Let us focus on the research question (1) in parts a to e.*

a\. Let $\theta$ be the proportion of women among those who attempt suicide in Shandong, China. State $H_0$ and $H_1$ in terms of $\theta$. Is it a right-sided test, left-sided test, or two-sided test?

b\. Provide a graphic which summarize the relevant data.

c\. Report the sample proportion of women among those who attempted suicide in the data, and Z statistic, and p-value.

d\. At a significance level $\alpha = 0.05$, do you conclude $H_0$ or $H_1$?

e\. State the conclusion in the context.

*Let us focus on the research question (2) in the following parts.*

f\. How would you define $\theta$ in the context of research question (2)?

g\. State $H_0$ and $H_1$ in terms of $\theta$. Is it a right-sided test, left-sided test, or two-sided test?

h\. Provide a graphic which summarize the relevant data.

i\. Report the sample proportion of women among those who attempted suicide in the data, and Z statistic, and p-value.

j\. At a significance level $\alpha = 0.05$, do you conclude $H_0$ or $H_1$?

k\. State the conclusion in the context.

**References:**

Klonsky, E. D., May, A. M., & Saffer, B. Y. (2016). Suicide, Suicide Attempts, and Suicidal Ideation. Annual review of clinical psychology, 12, 307–330. <https://doi.org/10.1146/annurev-clinpsy-021815-093204>

Sun, J., Guo, X., Zhang, J., Wang, M., Jia, C., & Xu, A. (2015). Incidence and fatality of serious suicide attempts in a predominantly rural population in Shandong, China: a public health surveillance study. BMJ open, 5(2), e006762. <https://doi.org/10.1136/bmjopen-2014-006762>

