---
title: "Introduction to Probability Theory I"
format: 
  live-html:
    webr:
      packages: 
        - ggformula
        - mosaic
        - tidyverse
      repos:
        - https://projectmosaic.r-universe.dev
      resources: 
      cell-options:
        warning: true
        completion: false
execute: 
  warning: false 
  message: false
  echo: true
embed-resources: true
code-copy: true
code-overflow: wrap
toc: true
number-sections: true
css: ../styles.css
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}   
{{< include ../_extensions/r-wasm/live/_gradethis.qmd >}}

```{webr}
#| label: theme
#| include: false
#| edit: false
library(ggplot2)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

```{r}
#| label: setup-checkdown
#| include: false
library(checkdown)
library(ggformula)
library(mosaic)
library(tidyverse)
```

## Sample Space

In an experiment, and the set of all outcomes under consideration is referred to as the **sample space**.

*Example:* When we toss a coin, the sample space is $\Omega = \{ h, t \}$ where $h$ denotes an outcome of head and $t$ denotes an outcome of tail.

*Example:* In a random experiment\* of tossing a six-sided die, the sample space is $\Omega = \{ 1, 2, 3, 4, 5, 6 \}$, the set of six possible outcomes.

*Example:* In a random experiment\* of tossing a pair of six-sided dice, the sample space is $\Omega = \{ (1, 1), (1, 2), \dots, (6, 6) \}$, the set of 36 possible paired outcomes.

::: callout-note
## Definition of Experiment

We will see the term "experiment" used in two different ways in this course. In the context of probability theory, *experiment* refers to the probability process by which data is generated. Once we learn more about the practice of statistics, *experiment* refers to a data collection method in which an individual is randomly assigned to a treatment group.
:::

#### Group Discussion

What are some sample spaces we can define around characteristics of your classmates?'

## Probability Model

Given a sample space $\Omega$, a **probability model** is a function $P$: $\Omega \rightarrow \mathcal{R}$ which satisfies the two conditions:

1.  $P(s) \ge 0$ for any $s \in \Omega$ .
2.  $\sum_{ s \in \Omega } P(s) = 1$.

*Example:* Given $\Omega = \{ h, t \}$ for tossing a coin, assume $P(h) = 0.8$ and $P(t) = 0.2$. It is a valid probability model because $P(h) \ge 0$ and $P(t) \ge 0$, and

$$\sum_{ s \in \Omega } P(s) = P(t) + P(h) = 0.8 + 0.2 = 1 \, .$$

In this case, we say that "the probability of $h$ is 0.8" and "the probability of $t$ is 0.2." Remember that a **probability model** is a function which satisfies the two conditions, and the **probability** of an outcome is a number between zero and one which quantifies the likelihood of the outcome.

*Example:* Given $\Omega = \{ 1, 2, 3, 4, 5, 6 \}$ for tossing a six-sided die, assume all six outcomes are equally likely. The probability model which expresses this assumption can be written as $P(s) = \frac{1}{6}$ for $s = 1, 2, \dots, 6$. This probability model can be graphically represented as follows.

```{webr}
#| label: bar-plot
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 5
#| out-height: "40%" 
#| out-width: "40%"
#| fig-alt: "Probability model of tossing a die"

library(ggformula)
omega <- c( 1, 2, 3, 4, 5, 6 )
prob <- c( 1/6, 1/6, 1/6, 1/6, 1/6, 1/6 )
gf_col(prob ~ factor(omega),
         title = "Probability Model of Tossing A Die",
         xlab = "Outcome",
         ylab = "Probability") +
  ylim(0, 0.5)
```

## Event

Let $\Omega$ be a sample space. An **event** is a subset of $\Omega$.

*Example:* Given $\Omega = \{ 1, 2, 3, 4, 5, 6 \}$ for tossing a six-sided die, $O = \{ 1, 3, 5 \}$ and $E = \{ 2, 4, 6 \}$ are events because $O \subseteq \Omega$ and $E \subseteq \Omega$ . If any outcome in an event occurred, we say that the event occurred. For example, if we toss a six-sided die and an outcome of 1 occurred, we say that the event $O$ occurred.

The probability of $O$ is calculated as follows:

$$P(O) = P(1) + P(3) + P(5) = \frac{1}{6} + \frac{1}{6} + \frac{1}{6} = 0.5 \, .$$

Similarly, the probability of $E$ is $P(E) = 0.5$. Why?

What do you notice about the relationship between $O$ and $E$ as they relate to the sample space?

At the end of semester, suppose students earn one of the following letter grades: $\Omega = \{ a, b, c, d, f \}$. Historically, 15%, 20%, 30%, 20%, and 15% of students earned $a$, $b$, $c$, $d$, and $f$, respectively. The following probability model reflects this history: $P(a) = P(f) = 0.15$, $P(b) = P(d) = 0.2$, and $P(c) = 0.3$. It can be graphically expressed as follows.

```{webr}
#| label: bar-plot-grades
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 5
#| out-height: "40%" 
#| out-width: "40%"
#| fig-alt: "Probability model of letter grades"

omega <- c( "a", "b", "c", "d", "f" )
prob <- c( 0.15, 0.2, 0.3, 0.2, 0.15 )
gf_col(prob ~ omega, 
         title = "Probability Model of Letter Grades",
         xlab = "Letter Grade",
         ylab = "Probability",
         col = "white", 
         ylim = c(0, 0.5) )
```

Let $G$ be the event of passing this course by earning $a$, $b$, or $c$. The probability of $G = \{ a, b, c \}$ is

$$P(G) = P(a) + P(b) + P(c) = 0.15 + 0.2 + 0.3 = 0.65 \, .$$ Let $G^c$ be the event of not passing this course. In this case, $G^c = \{ d, f \}$, and we $G^c$ is called the **complementary** of $G$. We can calculate the probability of $G^c$ as

$$ P(G^c) = P(d) + P(f) = 0.2 + 0.15 = 0.35 \, . $$

Alternatively,

$$P(G^c) = 1 - P(G) = 1 - 0.65 = 0.35 \, .$$

**Note:** By convention, $P( \emptyset ) = 0$ where $\emptyset$ is the empty set. In words, the probability of no outcomes is zero.



## Independence

**Definition:** If two events $A$ and $B$ are **independent** then $P(A \cap B) = P(A)P(B)$. In other words, $A$ and $B$ are independent when the probability of $B$ does not change the probability of $A$.

**Definition:** We say two events $A$ and $B$ are **dependent** when $P(A \cap B) \ne P(A)P(B)$. In other words, $A$ and $B$ are dependent when the condition of $B$ changes the probability of $A$.

**Example:** Consider the experiment of tossing a coin three times, write out the sample space.  


a. Assume all outcomes are equally likely. Write out or graph the probability function.


b. Suppose we need two or more heads to win a game, write out this event as 
$W =$


b. What is the probability of $W$?




d. Now consider the event that $h$ occurred on the first toss, call the event $H_1$ and write out the possible outcomes of the event.



e. What is the probability of $H_1$ occurring?



f. What is the probability of both $W$ and $H_1$ occurring?



g. Are $W$ and $H_1$ independent or dependent? How do you know?


h. Now consider the event that $t$ occurred on the second toss, call the event $T_2$ and write out the possible outcomes of the event.


i. What is the probability of $T_2$ occurring?



j. What is the probability of both $W$ and $T_2$ occurring?



k. Are $W$ and $T_2$ independent or dependent? How do you know?



## Wrap-Up

In your groups work on the following problems. You can check your answers below. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Be ready to show your work on the board.

#### Problem 1.

Fill in the blanks.

a\. The set of all outcomes under consideration is called the \[s... s...\]

```{r}
#| label: wrap-1-a
#| echo: false
checkdown::check_question(answer = "sample space",
          button_label = "Check Your Answer")
```

b\. A group of outcomes is called an \[e...\].

```{r}
#| label: wrap-1-b
#| echo: false
checkdown::check_question(answer = "event",
          button_label = "Check Your Answer")
```

c\. A probability model is a \[f...\] which receives an outcome and returns a non-negative number.

```{r}
#| label: wrap-1-c
#| echo: false
checkdown::check_question(answer = "function",
          button_label = "Check Your Answer")
```

d\. If $P(s) = t$, the number $t$ is called the \[p...\] of an outcome $s$.

```{r}
#| label: wrap-1-d
#| echo: false
checkdown::check_question(answer = "probability",
          button_label = "Check Your Answer")
```

e\. The sum of the probabilities of all outcomes under consideration is equal to \[o...\].

```{r}
#| label: wrap-1-e
#| echo: false
checkdown::check_question(answer = "one",
          button_label = "Check Your Answer")
```

#### Problem 2

True or false?

a\. An experiment must be repeated many times in order to observe the same outcome each time.

```{r}
#| label: wrap-2-a
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

b\. Let $E = \{ a, b, c \}$. We say the event $E$ occurred when **all** outcomes $a$, $b$, and $c$ occurred.

```{r}
#| label: wrap-2-b
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

c\. The probability of an event $A$ can be below zero if an outcome in $A$ is unlikely.

```{r}
#| label: wrap-2-c
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

d\. The probability of an event $B$ can be exceed one if there are a large number of outcomes in $B$ and all outcomes are highly likely.

```{r}
#| label: wrap-2-d
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

e\. The probability of an event is always between zero and one inclusive, $[0,1]$.

```{r}
#| label: wrap-2-e
#| echo: false
checkdown::check_question(answer = "True",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

#### Problem 3

Let $\Omega = \{ h, t \}$. Calculate $P(h)$ under the given assumption.

a\. $h$ and $t$ are equally likely.

```{r}
#| label: wrap-3-a
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

b\. The probability of $h$ is 9 times of the probability of $t$.

```{r}
#| label: wrap-3-b
#| echo: false
checkdown::check_question(answer = 0.9,
          button_label = "Check Your Answer")
```

c\. $h$ is impossible.

```{r}
#| label: wrap-3-c
#| echo: false
checkdown::check_question(answer = 0,
          button_label = "Check Your Answer")
```

d\. $t$ is impossible.

```{r}
#| label: wrap-3-d
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

#### Problem 4

Let $\Omega = \{ 1, 2, 3, 4 \}$ and $P$ be the probability model such that $P(s) = c \cdot s$ for $s = 1, 2, 3, 4$. Calculate the following**.**

a\. $c$

```{r}
#| label: wrap-4-a
#| echo: false
checkdown::check_question(answer = 0.1,
          button_label = "Check Your Answer")
```

b\. $P(A)$ where $A = \{ 1, 3, 4 \}$

```{r}
#| label: wrap-4-b
#| echo: false
checkdown::check_question(answer = 0.8,
          button_label = "Check Your Answer")
```

## Assignment

Check your answer by inputting them below their respective question. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 instead of 1/2) unless specified.

Though you are checking the numeric answers here, you should make sure your written work is complete and your mathematical notation is correct as both will be graded on the homework quiz.

#### Problem 1

Let $\Omega = \{ a, b, c, d \}$ be the sample space. Assume $a$ and $d$ are equally likely outcomes, $b$ and $c$ are equally likely outcomes, and the likelihood of $b$ is four times of $a$. Calculate the following.

a\. $P(a)$

```{r}
#| label: asn-1-a
#| echo: false
checkdown::check_question(answer = 0.1,
          button_label = "Check Your Answer")
```

b\. $P(b)$

```{r}
#| label: asn-1-b
#| echo: false
checkdown::check_question(answer = 0.4,
          button_label = "Check Your Answer")
```

c\. $P(c)$

```{r}
#| label: asn-1-c
#| echo: false
checkdown::check_question(answer = 0.4,
          button_label = "Check Your Answer")
```

d\. $P(d)$

```{r}
#| label: asn-1-d
#| echo: false
checkdown::check_question(answer = 0.1,
          button_label = "Check Your Answer")
```

e\. $P(a) + P(b) + P(c) + P(d)$

```{r}
#| label: asn-1-e
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

#### Problem 2

Suppose a computer randomly generates a lowercase alphabet (i.e., $\Omega = \{ a, b, \dots, z \}$ ). Let $V$ be the event of a vowel. Let $C$ be the event of a consonant. Assume all vowels are equally likely to be generated, all consonants are equally likely to be generated, and the likelihood of a vowel is twice of a consonant. Calculate the following.

a\. $P(C)$

```{r}
#| label: asn-2-a
#| echo: false
checkdown::check_question(answer = 0.6774,
          button_label = "Check Your Answer")
```

b\. $P(V)$

```{r}
#| label: asn-2-b
#| echo: false
checkdown::check_question(answer = 0.3226,
          button_label = "Check Your Answer")
```

c\. $P(V^c)$

```{r}
#| label: asn-2-c
#| echo: false
checkdown::check_question(answer = 0.6774,
          button_label = "Check Your Answer")
```

d\. $P(C \cup V)$

```{r}
#| label: asn-2-d
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

e\. $P(C \cap V)$

```{r}
#| label: asn-2-e
#| echo: false
checkdown::check_question(answer = 0,
          button_label = "Check Your Answer")
```

#### Problem 3

Let $\Omega = \{ a, b, c \}$ be the sample space. Let $A = \{ a, b \}$ and $B = \{ b, c \}$. Calculate $P(A \cap B)$ under the following assumption.

a\. All outcomes in $\Omega$ are equally likely.

```{r}
#| label: asn-3-a
#| echo: false
checkdown::check_question(answer = 0.3333,
          button_label = "Check Your Answer")
```

b\. $P(A) = 0.7$ and $P(B) = 0.7$

```{r}
#| label: asn-3-b
#| echo: false
checkdown::check_question(answer = 0.4,
          button_label = "Check Your Answer")
```

c\. $P(b) = 2 \cdot P(a)$ and $P(c) = 2 \cdot P(b)$

```{r}
#| label: asn-3-c
#| echo: false
checkdown::check_question(answer = 0.2857,
          button_label = "Check Your Answer")
```


