---
title: "Bayes Theorem"
format: 
  live-html:
    webr:
      packages: 
        - ggformula
        - mosaic
        - tidyverse
      repos:
        - https://projectmosaic.r-universe.dev
      resources: 
        - https://raw.githubusercontent.com/jcanner/math-320/refs/heads/master/note-outlines/3-intro-probability-ii/data/titanic.csv 
      cell-options:
        warning: true
        completion: false
execute: 
  warning: false 
  message: false
  echo: true
embed-resources: true
code-copy: true
code-overflow: wrap
toc: true
number-sections: true
css: ../styles.css
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}\
{{< include ../_extensions/r-wasm/live/_gradethis.qmd >}}

```{webr}
#| label: theme
#| include: false
#| edit: false
library(ggplot2)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

```{r}
#| label: setup-checkdown
#| include: false
library(checkdown)
library(ggformula)
library(mosaic)
library(tidyverse)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

## Bayes' Theorem

Some events are observable, while some are not. For example, we can observe whether a lie detector indicates positive or negative, but we never know whether the subject lied or not. Suppose a criminal suspect is under investigation. Let $B$ denote the event that the suspect lies (unknown) and $A$ denote the event that the lie detector indicates a lie (observable). Assume the following.

1.  If a suspect lies, the lie detector indicates that it is a lie with a probability of 0.8.
2.  If a suspect tells truth, the lie detector indicates that it is a lie with a probability of 0.05.
3.  A suspects lies with a probability of 0.9.

Given the notations $A$ and $B$, the above assumptions can be expressed as

- $P(A \mid B) = 0.8$,

- $P(A \mid B^c) = 0.05$, and

- $P(B) = 0.9$.

If the investigator observes a lie signal, what is the probability that the suspect lied? In other words, what is $P(B \mid A)$?

**Theorem:** For two events $A$ and $B$, the conditional probability of $B$ given $A$ is

$$P(B \mid A) = \frac{ P(A \mid B) }{ P(A) }$$

provided that $P(A) > 0$. This statement is known as the **Bayes' Theorem**.

**Note:** Since $B$ and $B^c$ are a partition of any sample space,

$$P(A) = P(A \mid B) \cdot P(B) + P(A \mid B^c) \cdot P(B^c) \, . $$ Therefore, the Bayes' Theorem can be also written as

$$P(B \mid A) = \frac{ P(A \mid B) \cdot P(B) }{ P(A \mid B) \cdot P(B) + P(A \mid B^c) \cdot P(B^c) } \, .$$ {#eq-bayes}

**Note:** Recall that $P(E^c) = 1 - P(E)$ for any event $E$. This property holds for a conditional probability. In other words, $P(E^c \mid C) = 1 - P(E \mid C)$ for any possible condition $C$. Therefore, @eq-bayes can be also written as

$$ P(B \mid A) = \frac{ P(A \mid B) \cdot P(B) }{ P(A \mid B) \cdot P(B) + [1 - P(A^c \mid B^c)] \cdot [1 - P(B)] } \, . $$

**Example:** Let us revisit the lie detector scenario. Under the assumptions of $P(A \mid B) = 0.8$, $P(A \mid B^c) = 0.05$, and $P(B) = 0.9$, we can apply the Bayes' Theorem (in the form of @eq-bayes) as follows:

$$ P(B \mid A) = \frac{ (0.8)(0.9) }{ (0.8)(0.9) + (0.05)(1-0.9) } = 0.993 \, . $$

In words, if the lie detector indicated that the suspect lied, there is a 0.993 probability that the suspect really lied.

In the context of the lie detector scenario, the conditional probability $P(A \mid B)$ is called the **sensitivity**. A high sensitivity implies a high chance of detecting a lie.

The conditional probability $P(A^c \mid B^c)$ is called the **specificity**. A high specificity implies a high chance of detecting a truth.

The probability $P(B)$ is called the **prevalence**, and the conditional probability $P(B \mid A)$ is called the **positive predictive value**. A high prevalence implies that subjects lie frequently. @eq-bayes informs that the reliability of a lie detector depends on the sensitivity and specificity (how good the lie detector is), but it also depends on the prevalence (how untruthful or truthful subjects are). Given the sensitivity ($\texttt{sens}$) and specificity ($\texttt{spec}$), the positive predictive value ($\texttt{ppv}$) is a function of the prevalence ($\texttt{ppv}$) as follows:

$$\texttt{ppv} = \frac{ \texttt{sens} \cdot \texttt{prev} }{ \texttt{sens} \cdot \texttt{prev} + (1 - \texttt{spec}) \cdot (1 - \texttt{prev}) } \, .$$

The below R code demonstrates the relationship between $\texttt{ppv}$ and $\texttt{prev}$ at $\texttt{sens} = 0.8$ and $\texttt{spec} = 0.95$. The graph demonstrates that the positive predictive value $P(B \mid A)$ increases with respect to the prevalence $P(B)$. In other words, we need frequent lies (i.e., $P(B)$ is high) in order for the lie detector predicts lies well (i.e., $P(B \mid A)$ is high).

```{r, fig.height = 5, fig.width = 5, fig.alt = "The plot of positive predictive value as a function of prevalence."}
### input: sens, spec, prev
### output: ppv

pos_pred_val <- function( sens, spec, prev ) {
   num = sens * prev
   den = sens * prev + ( 1 - spec ) * ( 1 - prev )
   ppv = num / den
   ppv }

pos_pred_val( sens = 0.8, spec = 0.95, prev = 0.1 ) ### returns ppv at given assumption
pos_pred_val( sens = 0.8, spec = 0.95, prev = 0.001 ) ### returns ppv at given assumption

est_prev <- seq( 0, 1, 0.001 ) ### prevalence from 0 to 1 with an increment of 0.001
est_ppv <- pos_pred_val( sens = 0.8, spec = 0.95, prev = est_prev ) ### evaluate the function at the points

gf_line( est_ppv ~ est_prev, 
      title = "PPV vs. Prevalence",
      xlab = "Prevalence", ylab = "PPV" ) ### plot
```

**Example:** Let us continue from the previous example. If the lie detect is quiet (no indication of a lie), what is the probability that the suspect really did not lie? In other words, what is the conditional probability of $B^c$ given $A^c$? Applying the Bayes' Theorem.

$$\begin{aligned}
P(B^c \mid A^c) & = \frac{P(A^c \mid B^c) \cdot P(B^c)}{P(A^c)} \\
& = \frac{P(A^c \mid B^c) \cdot P(B^c)}{P(A^c \mid B^c) \cdot P(B^c) + P(A^c \mid B) \cdot P(B)} \\
& = \frac{(0.95)(0.1)}{(0.95)(1 - 0.9) + (1-0.8)(0.9)} \\
& = 0.345 \, .
\end{aligned}$$

Such a low probability implies that we cannot trust the suspect when there is no indication of a lie by the lie detector. It is only useful when the lie detector indicates that it is a lie.


## Prenatal Testing for Rare Diseases

::: {.callout-note}
## Context
These are original, hypothetical exercises inspired by the article's discussion of prenatal genetic screening and false positives for rare conditions. A screening result estimates risk; it does not diagnose a condition. The numbers below are chosen for practice, not for medical decision-making.
:::

### Bayes' Theorem and Screening

Let \(C\) denote “the condition is present” and \(+\) denote “the screening test is positive.” Then

$$
P(C \mid +) =
\frac{P(+ \mid C)P(C)}
{P(+ \mid C)P(C) + P(+ \mid \neg C)P(\neg C)}.
$$

### Rare-condition screening

A condition occurs in 1 out of every 2,000 pregnancies. A screening test detects the condition 95% of the time when it is present and has a 1% false-positive rate.

If a patient receives a positive result, what is the probability that the fetus actually has the condition?

::: {.callout-tip collapse="true"}
## 1. Rare-condition screening

\[
P(C \mid +) =
\frac{(.95)(1/2000)}{(.95)(1/2000) + (.01)(1999/2000)}
\approx 0.0454.
\]

The probability is approximately **4.54%**.
:::

### Natural-frequency version

Use the information from Problem 1, but imagine 200,000 pregnancies.

- How many fetuses are expected to have the condition?
- How many true positives will the test produce?
- How many false positives will it produce?
- Of all positive results, what percentage are true positives?

::: {.callout-tip collapse="true"}
## 2. Natural-frequency version

Of 200,000 pregnancies, 100 are expected to have the condition. The test identifies \(0.95(100) = 95\) true positives.

Of the 199,900 pregnancies without the condition, 1% produce false positives: \(0.01(199,900) = 1,999\).

Thus, there are \(95 + 1,999 = 2,094\) positive results, of which

\[
\frac{95}{2,094} \approx 4.54\%
\]

are true positives.
:::

### Comparing two conditions

Test A screens for Condition A, which occurs in 1 in 100 pregnancies. Test B screens for Condition B, which occurs in 1 in 2,000 pregnancies. Both tests have 99% sensitivity and a 1% false-positive rate.

For each test, calculate \(P(\text{condition} \mid \text{positive test})\). Why can equally accurate tests lead to very different interpretations of a positive result?


::: {.callout-tip collapse="true"}
## 3. Comparing two conditions

For Condition A:

\[
\frac{(.99)(.01)}{(.99)(.01) + (.01)(.99)} = 0.50.
\]

So a positive test gives a **50%** probability that Condition A is present.

For Condition B:

\[
\frac{(.99)(.0005)}{(.99)(.0005) + (.01)(.9995)} \approx 0.0472.
\]

So a positive test gives only about a **4.72%** probability that Condition B is present. The key difference is the base rate: Condition B is much rarer.
:::


### Different risk group

In a higher-risk group, a condition occurs in 1 in 250 pregnancies. The same screening test from Problem 1 is used: 95% sensitivity and a 1% false-positive rate.

What is the probability of the condition after a positive result? Compare it with your answer to Problem 1. What changed, and what stayed the same?

::: {.callout-tip collapse="true"}
## 4. Different risk group

\[
\frac{(.95)(.004)}{(.95)(.004) + (.01)(.996)}
\approx 0.276.
\]

The probability after a positive result is about **27.6%**. The test's sensitivity and false-positive rate stayed the same; the prior probability increased.
:::

### A negative result

A condition occurs in 1 in 500 pregnancies. A screening test has 98% sensitivity and a 2% false-positive rate.

If the result is negative, what is the probability that the fetus still has the condition?

::: {.callout-tip collapse="true"}
## 5. A negative result

The false-negative rate is \(1 - .98 = .02\). Therefore,

\[
P(C \mid -) =
\frac{(.02)(.002)}{(.02)(.002) + (.98)(.998)}
\approx 0.0000409.
\]

The probability is approximately **0.0041%**, or about 1 in 24,450.
:::

### An advertising claim

A company advertises that its screening test is “99% accurate.” Suppose the condition occurs in 1 in 10,000 pregnancies, sensitivity is 99%, and specificity is 99%.

- Find the positive predictive value.  
- Is “99% accurate” enough information to tell a patient the chance that a positive result is correct? Explain using Bayes' theorem.  

::: {.callout-tip collapse="true"}
## 6. An advertising claim

\[
P(C \mid +) =
\frac{(.99)(.0001)}{(.99)(.0001) + (.01)(.9999)}
\approx 0.0098.
\]

The positive predictive value is about **0.98%**. No: a general accuracy claim omits the condition's prevalence, as well as the separate sensitivity and specificity values that Bayes' theorem requires.
:::

### A follow-up diagnostic test

A screening test yields a 4% probability that a fetus has a condition after a positive result. A follow-up diagnostic test has 99.9% sensitivity and 99.9% specificity.

Assuming the diagnostic test is positive, use Bayes' theorem to estimate the updated probability that the fetus has the condition. Explain why follow-up testing matters.

::: {.callout-tip collapse="true"}
## 7. A follow-up diagnostic test

Treat the 4% result as the new prior probability:

\[
P(C \mid +) =
\frac{(.999)(.04)}{(.999)(.04) + (.001)(.96)}
\approx 0.9765.
\]

The updated probability is approximately **97.65%**. A highly accurate follow-up test can substantially reduce uncertainty after a screening result.
:::


#### Source context

The problems are based on the general distinction between prenatal screening and diagnostic testing discussed in [the referenced New York Times article](https://www.nytimes.com/2022/01/01/upshot/pregnancy-birth-genetic-testing.html). For current clinical guidance, consult a qualified health-care professional.


## Bayes Theorem and Machine Learning

### Example 1: A Naive Bayes spam classifier

A Naive Bayes classifier uses Bayes' theorem to estimate the probability that a message belongs to a class, such as *spam* or *not spam*, after observing its words. The word *naive* refers to an assumption: once the class is known, the model treats the observed word features as independent.

Suppose that, in a training set:

- 20% of emails are spam: \(P(S) = 0.20\);
- 75% of spam emails contain the word “winner”: \(P(W \mid S) = 0.75\);
- 5% of non-spam emails contain “winner”: \(P(W \mid \neg S) = 0.05\).

For an email that contains “winner,” the classifier calculates

\[
P(S \mid W) =
\frac{P(W \mid S)P(S)}
{P(W \mid S)P(S) + P(W \mid \neg S)P(\neg S)}.
\]

\[
P(S \mid W) =
\frac{(.75)(.20)}{(.75)(.20) + (.05)(.80)}
= \frac{.15}{.19}
\approx 0.789.
\]

So the model estimates a **78.9% probability** that the email is spam. In practice, it combines evidence from many words:

\[
P(S \mid w_1, \ldots, w_n) \propto P(S)\prod_{i=1}^{n}P(w_i \mid S).
\]

The class with the larger posterior probability is the prediction. This is the same update as in the screening problems: begin with a prior probability, then revise it using evidence.

### Example 2: Bayesian A/B testing

In a Bayesian A/B test, we treat each version's conversion rate as an unknown probability and update our beliefs as observations arrive. Unlike a classical A/B test, the central result can be stated directly as a probability about the two versions.

Suppose a website runs two versions of a sign-up page. Before collecting data, use a uniform prior for each conversion rate:

\[
p_A, p_B \sim \operatorname{Beta}(1,1).
\]

After the experiment:

| Version | Visitors | Sign-ups | Non-sign-ups |
|---|---:|---:|---:|
| A | 100 | 12 | 88 |
| B | 100 | 18 | 82 |

With a Beta prior and binomial conversion data, the posterior distribution is

\[
p \mid \text{data} \sim \operatorname{Beta}(\alpha + \text{sign-ups},\; \beta + \text{non-sign-ups}).
\]

Thus,

\[
p_A \mid \text{data} \sim \operatorname{Beta}(13,89),
\qquad
p_B \mid \text{data} \sim \operatorname{Beta}(19,83).
\]

The posterior means are

\[
E[p_A \mid \text{data}] = \frac{13}{102} \approx 12.7\%,
\qquad
E[p_B \mid \text{data}] = \frac{19}{102} \approx 18.6\%.
\]

To make a decision, draw many samples from the two posterior distributions and count how often \(p_B > p_A\). That fraction estimates \(P(p_B > p_A \mid \text{data})\): the probability, given the model and observed data, that Version B has the higher conversion rate. This illustrates Bayes' theorem as a learning process—prior beliefs are updated to posterior beliefs as data accumulate.

## Warm-up

*You must input your answers in the [Google form](https://docs.google.com/forms/d/e/1FAIpQLScERQAIiFLzCohcg3cFH8HshlZ2eVvQutX6WOMZOZa0kZdtJg/viewform?usp=dialog). All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Any typo will result in point deduction in the auto-grading system. Any submission after the due date specified in Canvas will not be graded.*

**Problem 1.** Assume $P(A \mid B) = 0.9$, $P(A \mid B^c) = 0.3$, and $P(B) = 0.7$. Calculate the following probabilities.

a\. $P(B^c)$

b\. $P(A)$

c\. $P(A^c)$

d\. $P(A^c \mid B)$

e\. $P(A^c \mid B^c)$

f\. $P(B \mid A)$

g\. $P(B^c \mid A)$

h\. $P(B^c \mid A^c)$

i\. $P(B \mid A^c)$

**Problem 2.** Suppose any news is good or bad. If there is good news in the morning, there is good news in the evening with a probability of 0.75. If there is bad news in the morning, there is bad news in the evening with a probability of 0.25. Assume morning news is good 51% of the time. Suppose we missed morning news today, and there is good news in the evening. What is the probability that today's morning was good?

**Problem 3.** Suppose we toss a coin inside a room. If it rains outside, head occurs with a probability of 0.99. If it does not rain outside, head occurs with a probability of 0.1. Suppose it rains 10% of the time. Given head, calculate the probability that it is raining outside.

**Problem 4.** Suppose we toss a coin inside a room. If it rains outside, head occurs with a probability of 0.99. If it does not rain outside, head occurs with a probability of 0.1. Suppose it rains 10% of the time. Given head, calculate the probability that it is raining outside.

## Assignment

*You must input your answers in the [Google form](https://docs.google.com/forms/d/e/1FAIpQLSfSTxrK_g-dmNXoRnAvlogHcSs6l7mZ3KPyC07r_BRGLVj7oA/viewform?usp=dialog). All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Any typo will result in point deduction in the auto-grading system. Any submission after the due date specified in Canvas will not be graded.*

**Problem 1.** Assume $P(A) = 0.5$, $P(B \mid A) = 0.8$, and $P(B^c \mid A^c) = 0.7$. Calculate the following probabilities.

a\. $P(A^c)$

b\. $P(B)$

c\. $P(B^c)$

d\. $P(B \mid A^c)$

e\. $P(B^c \mid A^c)$

f\. $P(A \mid B)$

g\. $P(A^c \mid B)$

h\. $P(A \mid B^c)$

i\. $P(A^c \mid B^c)$

j\. $P(A \mid A^c)$

k\. $P(A \mid A)$

**Problem 2.** In an airport, a body scanner beeps when it detects a metal (a potential concealed weapon). If a traveler enters the body scanner with a metal, it beeps with a probability 0.999. If a traveler enters the body scanner without a metal, it beeps (false alarm) with a probability of 0.1.

a\. Assume 0.01% of travelers (1 in 10,000) enter the body scanner with a metal. If a body scanner beeps, what is the probability that the traveler actually has a metal?

b\. Repeat part a assuming 10% of travelers enter the body scanner with a metal.

**Problem 3.** For fair criminal trials, defendants are assumed to be innocent until proven guilty, and prosecutors must provide sufficient evidence in order to conclude that defendants are guilty. Assume the following:

- If a defendant is innocent, it is concluded that the defendant is guilty with a probability of 0.05.

- If a defendant is guilty, it is concluded that the defendant is guilty with a probability of 0.9.

As a hypothetical scenario, suppose prosecutors accuse any people so that 99% of defendants are innocent. If the conclusion is guilty, what is the probability that the defendant is truly guilty?

**Problem 4.** Until 2013, all employees in Canadian workplaces were subject to drug test by a random selection.^1^ The drug test was intended to identify subjects who use illegal drugs, but it had been criticized that too many innocent employees were tested. In 2013, the Supreme Court of Canada ruled that there must be a reasonable cause (e.g., suspicious actions) to drug test an employee. This new policy would narrow the target population of drug test from all employees to suspicious employees. Let $U$ denote a drug use in a target population and $T$ denote a postive drug test result. Assume both sensitivity and specificity of the drug test have been constant at 0.95, and assume that only 5% are drug users among all employees and 95% are drug users among those who act suspiciously. Calculate the following.

a\. $P(T \mid U)$

b\. $P(T \mid U^c)$

c\. $P(T^c \mid U^c)$

d\. $P(T^c \mid U)$

e\. $P(U)$ under the old policy (prior to 2013)

f\. $P(U)$ under the new policy

g\. Under the [old]{.underline} policy, what is the probability that the subject is [not a drug user]{.underline} when the test result is [positive]{.underline}? (i.e., accuse an innocent subject)

h\. Under the [old]{.underline} policy, what is the probability that the subject is [a drug user]{.underline} when the test result is [negative]{.underline}? (i.e., free a guilty subject)

i\. Under the [new]{.underline} policy, what is the probability that the subject is [not a drug user]{.underline} when the test result is [positive]{.underline}?

j\. Under the [new]{.underline} policy, what is the probability that the subject is [a drug user]{.underline} when the test result is [negative]{.underline}?

k\. Multiple-choice: All employees had been targeted under the old policy, and only suspicious employees have been targeted under the new policy. If the test result is negative, the subject is concluded to be clean, and will not be tested for the next three years. If the test result is positive, the subject is temporarily concluded to be a potential risk in the workplace, and will be observed until proven to be clean. According to the probability calculations in previous parts, the \[...\] policy raises the concern of safety in the workplace, and the \[...\] policy raises the concern of privacy.

\(A\) old; old

\(B\) old; new

\(C\) new; old

\(D\) new; new

\(E\) both; both

^1^ Hurley, W. J. (2015). Mandatory drug testing in the Canadian workplace: A note on the recent supreme court decision and its effect on the misclassification rate of non-drug users, Chance, 28(1), 21-26. \url{http://chance.amstat.org/2015/02/drug-testing/}

## Assignment\*

*You must submit hard copy. All answers must be explained using both notation and words in complete sentences. Lack of explanation, misuse of notation, or incomplete sentences will result in partial credits only.*

**Problem 5.** Image you are the wife of a man who was in Titanic which sank in the morning of April 15, 1912. News on radio said 30% of all passengers survived, 25% were men among survivors, and 80% were men among non-survivors. You do not know whether your husband survived or not.

a\. What percent of passengers were men?

b\. What is the probability that your husband survived?

## Wrap-Up

In your groups work on the following problems. You can check your answers below. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Be ready to show your work on the board.

#### Problem 1

Let $A$ and $B$ be two events. Fill in the blanks.

a.  $P(A \cap B)$ is called the \[j... p...\] of $A$ and $B$.

```{r}
#| label: wrap-1-a
#| echo: false
checkdown::check_question(answer = "joint probability",
          button_label = "Check Your Answer")
```

b.  $P(A)$ is called the \[m... p...\] of $A$.

```{r}
#| label: wrap-1-b
#| echo: false
checkdown::check_question(answer = "marginal probability",
          button_label = "Check Your Answer")
```

c.  $P( B \mid A)$ is called the \[c... p...\] of the event $B$ given the possible condition $A$.

```{r}
#| label: wrap-1-c
#| echo: false
checkdown::check_question(answer = "conditional probability",
          button_label = "Check Your Answer")
```

#### Problem 2

Let $\Omega = \{ 1, 2, 3, 4, 5, 6 \}$ for tossing a six-sided die. Assume all six outcomes are equally likely. Let $A = \{ 1 \}$. Let $O$ be the event of observing an odd face. Let $E$ be the event of observing an even face. Calculate the following.

a.  $P(1)$

```{r}
#| label: wrap-2-a
#| echo: false
checkdown::check_question(answer = 0.1667,
          button_label = "Check Your Answer")
```

b.  $P(O)$

```{r}
#| label: wrap-2-b
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

c.  $P(A \mid O)$

```{r}
#| label: wrap-2-c
#| echo: false
checkdown::check_question(answer = 0.3333,
          button_label = "Check Your Answer")
```

d.  $P(A \mid E)$

```{r}
#| label: wrap-2-d
#| echo: false
checkdown::check_question(answer = 0,
          button_label = "Check Your Answer")
```

e.  $P(O \mid A )$

```{r}
#| label: wrap-2-e
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

#### Problem 3

Let $A$ and $B$ be two events. True or false?

a.  $P(A \mid B) = P(B \mid A)$

```{r}
#| label: wrap-3-a
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

b.  $P(A) \ge P(A \mid B)$

```{r}
#| label: wrap-3-b
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

c.  $P(B) \ge P(A \mid B)$

```{r}
#| label: wrap-3-c
#| echo: false
checkdown::check_question(answer = "False",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

d.  $P(A \mid B) \ge P(A \cap B)$

```{r}
#| label: wrap-3-d
#| echo: false
checkdown::check_question(answer = "True",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

e.  $P(A) \ge P(A \cap B)$

```{r}
#| label: wrap-3-e
#| echo: false
checkdown::check_question(answer = "True",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

f.  $P(A) + P(A^c) = 1$

```{r}
#| label: wrap-3-f
#| echo: false
checkdown::check_question(answer = "True",
                          options = c("True",
                                      "False"),
                          type = "select",
                          button_label = "Check Your Answer")
```

#### Problem 4

Let $\Omega$ be a sample space, and $A$, $B$, and $C$ be events. Fill in the blanks.

a.  If $A \cap B = \emptyset$, $A \cap C = \emptyset$, and $B \cap C = \emptyset$, we say that $A$, $B$, and $C$ are a \[p... d...\].

```{r}
#| label: wrap-4-a
#| echo: false
checkdown::check_question(answer = "pairwise disjoint",
          button_label = "Check Your Answer")
```

b.  If $A$, $B$, and $C$ are pairwise disjoint, and their union is $\Omega$, we say that $A$, $B$, and $C$ are a \[p...\] of $\Omega$.

```{r}
#| label: wrap-4-b
#| echo: false
checkdown::check_question(answer = "partition",
          button_label = "Check Your Answer")
```

c.  $A$ and $A^c$ are a \[p...\] of $\Omega$.

```{r}
#| label: wrap-4-c
#| echo: false
checkdown::check_question(answer = "partition",
          button_label = "Check Your Answer")
```

#### Problem 5

Let $\Omega$ be a sample space, and $A$, $B$, and $C$ are a partition of $\Omega$. Calculate the following.

a.  $P(A) + P(B) + P(C)$

```{r}
#| label: wrap-5-a
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

b.  $P(A) + P(A^c)$

```{r}
#| label: wrap-5-b
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

#### Problem 6

Suppose 47% of the population are female nonsmokers, 43% are male nonsmokers, 2% are female smokers, and 8% are male smokers. Suppose we randomly select a person from the population. Calculate:

a.  the probability of selecting a smoker

```{r}
#| label: wrap-6-a
#| echo: false
checkdown::check_question(answer = 0.1,
          button_label = "Check Your Answer")
```

b.  the probability of selecting a male

```{r}
#| label: wrap-6-b
#| echo: false
checkdown::check_question(answer = 0.51,
          button_label = "Check Your Answer")
```

#### Problem 7

It rains on Monday with a probability of 0.25. If it rains on Monday, it rains on Tuesday with a probability of 0.5. If it does not rain on Monday, it rains on Tuesday with a probability of 0.9. Calculate:

a.  the probability of raining on both Monday and Tuesday

```{r}
#| label: wrap-7-a
#| echo: false
checkdown::check_question(answer = 0.125,
          button_label = "Check Your Answer")
```

b.  the probability of raining on Tuesday

```{r}
#| label: wrap-7-b
#| echo: false
checkdown::check_question(answer = 0.8,
          button_label = "Check Your Answer")
```

#### Problem 8

Fill in the blank.

a.  The \[L... of T... P...\] is applied in Problems 6a and 6b.

```{r}
#| label: wrap-8-a
#| echo: false
checkdown::check_question(answer = "Law of Total Probability",
          button_label = "Check Your Answer")
```

b.  The \[M... R...\] is applied in Problems 7a.

```{r}
#| label: wrap-8-b
#| echo: false
checkdown::check_question(answer = "Multiplication Rule",
          button_label = "Check Your Answer")
```

c.  The \[w... a...\] of two conditional probabilities is used in Problems 7b.

```{r}
#| label: wrap-8-c
#| echo: false
checkdown::check_question(answer = "weighted average",
          button_label = "Check Your Answer")
```

#### Problem 9

It is estimated that 40%, 11%, 45%, and 4% of Americans have blood type A, B, O, and AB, respectively. It is estimated that 7%, 8%, 9%, and 7% of blood type A, B, O, and AB, respectively, develop a certain disease. Calculate the prevalence of the disease in the population. In other words, if we randomly select an American, what is the probability that the subject has the disease?

```{r}
#| label: wrap-9
#| echo: false
checkdown::check_question(answer = 0.0801,
          button_label = "Check Your Answer")
```

## Assignment

Check your answer by inputting them below their respective question. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified.

Though you are checking the numeric answers here, you should make sure your written work is complete and your mathematical notation is correct as both will be graded on the homework quiz.

#### Problem 1

Suppose a student must earn a high school grade point average (HS GPA) of 2.0 or above to apply to a university. The population of interest is all students who take Calculus I in the university. Let $G_i$ be the HS GPA group of $[i, i + 1)$ for $i = 0, 1, 2, 3, 4$ where $[a, b)$ is an interval between $a$ and $b$ including $a$ and excluding $b$. Let $A$, $B$, $C$, $D$, and $F$ be the groups of students who earned letter grades of A, B, C, D, and F in Calculus I, respectively. Answer the following question given the data in @tbl-gpa.

::: {#tbl-gpa}
| Event | $G_0$ | $G_1$ | $G_2$ | $G_3$ | $G_4$ | Total |
|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|
|  $A$  |   0   |   0   |  12   |  146  |  16   |  174  |
|  $B$  |   0   |   0   |  43   |  216  |   9   |  268  |
|  $C$  |   0   |   0   |  44   |  167  |   3   |  214  |
|  $D$  |   0   |   0   |  33   |  48   |   1   |  82   |
|  $F$  |   0   |   0   |  32   |  56   |   1   |  89   |
| Total |   0   |   0   |  164  |  633  |  30   |  827  |

: Contingency table of HS GPA and Calculus I grade (count)
:::

a.  Calculate the probability of earning an A in Calculus I.

```{r}
#| label: asn-1-a
#| echo: false
checkdown::check_question(answer = 0.2104,
          button_label = "Check Your Answer")
```

b.  If a student's HS GPA is between 2 and 2.99, calculate the probability of earning an A in Calculus I.

```{r}
#| label: asn-1-b
#| echo: false
checkdown::check_question(answer = 0.0732,
          button_label = "Check Your Answer")
```

c.  If a student's HS GPA is 4 or above, calculate the probability of earning an A in Calculus I.

```{r}
#| label: asn-1-c
#| echo: false
checkdown::check_question(answer = 0.5333,
          button_label = "Check Your Answer")
```

d.  Suppose a passing grade is A, B, or C. Calculate the probability of passing Calculus I.

```{r}
#| label: asn-1-d
#| echo: false
checkdown::check_question(answer = 0.7932,
          button_label = "Check Your Answer")
```

e.  If a student's HS GPA is between 2 and 2.99, calculate the probability of passing Calculus I.

```{r}
#| label: asn-1-e
#| echo: false
checkdown::check_question(answer = 0.6037,
          button_label = "Check Your Answer")
```

f.  If a student's HS GPA is 4 or above, calculate the probability of passing passing Calculus I.

```{r}
#| label: asn-1-f
#| echo: false
checkdown::check_question(answer = 0.9333,
          button_label = "Check Your Answer")
```

g.  If a student's HS GPA is 4 or above, calculate the probability of not passing passing Calculus I.

```{r}
#| label: asn-1-g
#| echo: false
checkdown::check_question(answer = 0.0667,
          button_label = "Check Your Answer")
```

h.  If a student passed Calculus I, calculate the probability that the student's HS GPA is 3 or above.

```{r}
#| label: asn-1-h
#| echo: false
checkdown::check_question(answer = 0.8491,
          button_label = "Check Your Answer")
```

#### Problem 2

Autism is a developmental disorder. Autistic children have difficulties with social interactions, verbal and nonverbal communications, and repetitive behaviors. According to the Centers for Disease Control and Prevention (CDC) in 2021, about one in 27 boys and one in 116 girls are identified with autism in the United States. It is estimated that the ratio of boys is about 1.05 in the United States. Let $A$ be autism and $B$ be a boy.

a.  Calculate $\frac{P( A \mid B )}{P( A \mid B^c )}$ (called the odds).

```{r}
#| label: asn-2-a
#| echo: false
checkdown::check_question(answer = 4.2963,
          button_label = "Check Your Answer")
```

b.  Interpret the ratio of the two conditional probabilities in the context.

<!-- According to the CDC estimate, the prevalence of autism among boys is about 4.3 times of -->

<!-- the prevalence of autism among girls -->

c.  @tbl-autism is a contingency table of 10,000 hypothetical children. Fill in each cell. Round to whole numbers.

::: {#tbl-autism}
| Event | $A$ | $A^c$ | Total  |
|:-----:|:---:|:-----:|:------:|
|  $B$  |  i  |  ii   |  iii   |
| $B^c$ | iv  |   v   |   vi   |
| Total | vii | viii  | 10,000 |

: Contingency table of autism and sex (count)
:::

c.i.

```{r}
#| label: asn-2-c-i
#| echo: false
checkdown::check_question(answer = 190,
          button_label = "Check Your Answer")
```

c.ii.

```{r}
#| label: asn-2-c-ii
#| echo: false
checkdown::check_question(answer = 4932,
          button_label = "Check Your Answer")
```

c.iii.

```{r}
#| label: asn-2-c-iii
#| echo: false
checkdown::check_question(answer = 5122,
          button_label = "Check Your Answer")
```

c.iv.

```{r}
#| label: asn-2-c-iv
#| echo: false
checkdown::check_question(answer = 42,
          button_label = "Check Your Answer")
```

c.v.

```{r}
#| label: asn-2-c-v
#| echo: false
checkdown::check_question(answer = 4836,
          button_label = "Check Your Answer")
```

c.vi.

```{r}
#| label: asn-2-c-vi
#| echo: false
checkdown::check_question(answer = 4878,
          button_label = "Check Your Answer")
```

c.vii.

```{r}
#| label: asn-2-c-vii
#| echo: false
checkdown::check_question(answer = 232,
          button_label = "Check Your Answer")
```

c.viii.

```{r}
#| label: asn-2-c-viii
#| echo: false
checkdown::check_question(answer = 9768,
          button_label = "Check Your Answer")
```

d.  If a child is autistic, calculate the probability that the child is a boy.

```{r}
#| label: asn-2-d
#| echo: false
checkdown::check_question(answer = 0.819,
          button_label = "Check Your Answer")
```

e\. Are $A$ and $B$ independent? Justify.

```{r}
#| label: asn-2-e
#| echo: false
checkdown::check_question(answer = "dependent",
                          options = c("independent",
                                      "dependent"),
                          type = "select",
                          button_label = "Check Your Answer")
```

#### Problem 3

Let $\Omega$ be a sample space. Let $A$ be an event, and suppose $C_1$, $C_2$, and $C_3$ are a partition of $\Omega$. Assume $P(A) = 0.5$. Calculate the following.

a.  $P(\Omega)$

```{r}
#| label: asn-3-a
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

b.  $P(A \cap C_1) + P(A \cap C_2) + P(A \cap C_3)$

```{r}
#| label: asn-3-b
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

c.  $P(C_1) + P(C_2) + P(C_3)$

```{r}
#| label: asn-3-c
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

d.  $P(A \mid C_1) \cdot P(C_1) + P(A \mid C_2) \cdot P(C_2) + P(A \mid C_3) \cdot P(C_3)$

```{r}
#| label: asn-3-d
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

#### Problem 4

Suppose 2% of CSUMB students are Mathematics Majors. Among Mathematics Majors, 30% are seniors. What percent of CSUMB students are mathematics seniors? (Do not include the percent sign when you input the numeric answer.)

```{r}
#| label: asn-4
#| echo: false
checkdown::check_question(answer = 0.6,
          button_label = "Check Your Answer")
```

#### Problem 5

Among vaccinated people, 25% have the flu. Among non-vaccinated people, 75% have the flu. What percent of people should be vaccinated to maintain a prevalence of 30% in the population? (Do not include the percent sign when you input the numeric answer.)

```{r}
#| label: asn-5
#| echo: false
checkdown::check_question(answer = 90,
          button_label = "Check Your Answer")
```

#### Problem 6

If a woman smokes during pregnancy, the probability of a premature birth is 0.1. If a woman does not smoke during pregnancy, the probability of a premature birth is 0.05. If 7 in 100 pregnant women smoke on average, what is the probability of a premature birth in the population? (A probability is a number between zero and one.)

```{r}
#| label: asn-6
#| echo: false
checkdown::check_question(answer = 0.0535,
          button_label = "Check Your Answer")
```

#### Problem 7

Titanic was a British passenger liner. It sank in the North Atlantic Ocean in the morning of April 15, 1912. The cause was collision with an iceberg, and many passengers lost lives. **titanic.csv** is a dataset which includes information about 1,316 passengers. There are five columns:

- **id**: passenger identification number

- **class**: passenger class (first, second, or third)

- **age**: passenger's age group (adult or child)

- **sex**: passenger's sex (female or male)

- **survived**: passenger's survival status (no or yes)

To use R, first load the packages needed for data analysis.

```{webr}
#| label: setup-webr
library(ggformula)  #graphics
library(mosaic)   #summary statistics
library(tidyverse) #data management
```

Then, we want to read in the data.

```{webr}
#| label: read-data
titanic <- read_csv( "titanic.csv",
                     show_col_types = FALSE) ### read data, hide messages about columns
head(titanic, 5) ### print the top 5 lines of the data
```

We can use two functions: `count()` and `props()` to evaluate our specified categories, `~ x`.

```{webr}
#| label: survival-count
counts( ~survived, data = titanic ) #counts
```

```{webr}
#| label: survival-props
props( ~survived, data = titanic) # proportions
```

We can use the formula `~ x | z` to calculate conditional counts and proportions of `x` conditioned on `z`.

```{webr}
#| label: survival-sex-demo

### count by groups
counts(~ survived |sex, data = titanic )

### proportions survived by groups
props(~ survived | sex, data = titanic)
```

The last R output can be interpreted as follows: about 27.5% of female passengers did not survive, about 72.5% of female passengers survived, about 79.9% of male passengers did not survive, and about 20.1% of male passengers survived.

**Use the Law of Total Probability and the Multiplication Rule to calculate the probability of survival in three different ways. Though the code helps make the calculations efficient, write out the mathematical notation each value and calculation represents.**

a.  Estimate the probability of passengers survived. Answer based on the partition by passenger class (first, second, and third).

```{webr}
#| label: survival-class
#| exercise: ex_class
#| caption: Edit the code to calculate a each probability by replacing x and z

### calculate the probability of each class
c <- props(~z, data = titanic)  #change this code
c #prints out answer

### calculate the probability of  survival given class
s_c <- props(~ x | z, data = titanic)  #change this code
s_c #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_c * select(c, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_c * select(c, -response))
```

::: {.solution exercise="ex_class"}
##### Solution

```{webr}
#| exercise: ex_class
#| solution: true
#| echo: false

### calculate the probability of each class
c <- props(~class, data = titanic)  #change this code
c #prints out answer

### calculate the probability of  survival given class
s_c <- props(~ survived | class, data = titanic)  #change this code
s_c #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_c * select(c, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_c * select(c, -response))
```
:::

```{webr}
#| exercise: ex_class
#| check: true
#| echo: false
gradethis::grade_this_code()
```

If you did the calculation by hand, you can check your answer here:

```{r}
#| label: asn-7-a
#| echo: false
checkdown::check_question(answer = 0.3792,
          button_label = "Check Your Answer")
```

b.  Estimate the probability of passengers survived. Answer based on the partition by age group (adult and child).

```{webr}
#| label: survival-age
#| exercise: ex_age
#| caption: Edit the code to calculate a each probability by replacing x and z

### calculate the probability of each age
a <- props(~z, data = titanic)  #change this code
a #prints out answer

### calculate the probability of  survival given age
s_a <- props(~ x | z, data = titanic)  #change this code
s_a #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_a * select(a, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_a * select(a, -response))
```

::: {.solution exercise="ex_age"}
##### Solution

```{webr}
#| exercise: ex_age
#| solution: true
#| echo: false

### calculate the probability of each age
a <- props(~age, data = titanic)  #change this code
a #prints out answer

### calculate the probability of  survival given age
s_a <- props(~ survived | age, data = titanic)  #change this code
s_a #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_a * select(a, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_a * select(a, -response))
```
:::

```{webr}
#| exercise: ex_age
#| check: true
#| echo: false
gradethis::grade_this_code()
```

If you did the calculation by hand, you can check your answer here:

```{r}
#| label: asn-7-b
#| echo: false
checkdown::check_question(answer = 0.3792,
          button_label = "Check Your Answer")
```

c.  Estimate the probability of passengers survived. Answer based on the partition by sex (female and male).

```{webr}
#| label: survival-sex
#| exercise: ex_sex
#| caption: Edit the code to calculate a each probability by replacing x and z

### calculate the probability of each sex
se <- props(~z, data = titanic)  #change this code
se #prints out answer

### calculate the probability of  survival given sex
s_se <- props(~ x | z, data = titanic)  #change this code
s_se #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_se * select(se, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_se * select(se, -response))
```

::: {.solution exercise="ex_sex"}
##### Solution

```{webr}
#| exercise: ex_sex
#| solution: true
#| echo: false

### calculate the probability of each sex
se <- props(~sex, data = titanic)  #change this code
se #prints out answer

### calculate the probability of  survival given sex
s_se <- props(~ survived | sex, data = titanic)  #change this code
s_se #prints out answer

## Do not adjust the code below ## 

# joint probabilities
s_se * select(se, -response) # removes the response column so element by element product is calculated

# marginal probability of survival
sum(s_se * select(se, -response))
```
:::

```{webr}
#| exercise: ex_sex
#| check: true
#| echo: false
gradethis::grade_this_code()
```

If you did the calculation by hand, you can check your answer here:

```{r}
#| label: asn-7-c
#| echo: false
checkdown::check_question(answer = 0.3792,
          button_label = "Check Your Answer")
```

d.  Explain why the three estimates in parts a, b, and c are the same.
