---
title: "Bayes Theorem"
format: 
  live-html:
    webr:
      packages: 
        - ggformula
        - mosaic
        - tidyverse
      repos:
        - https://projectmosaic.r-universe.dev
      resources: 
        - https://raw.githubusercontent.com/jcanner/math-320/refs/heads/master/note-outlines/3-intro-probability-ii/data/titanic.csv 
      cell-options:
        warning: true
        completion: false
execute: 
  warning: false 
  message: false
  echo: true
embed-resources: true
code-copy: true
code-overflow: wrap
toc: true
number-sections: true
css: ../styles.css
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}  
{{< include ../_extensions/r-wasm/live/_gradethis.qmd >}}

```{webr}
#| label: theme
#| include: false
#| edit: false
library(ggplot2)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

```{r}
#| label: setup-checkdown
#| include: false
library(checkdown)
library(ggformula)
library(mosaic)
library(tidyverse)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

## Bayes' Theorem

Some events are observable, while some are not. For example, we can observe whether a lie detector indicates positive or negative, but we never know whether the subject lied or not. Suppose a criminal suspect is under investigation. Let $L$ denote the event that the suspect lies (unknown) and $D$ denote the event that the lie detector indicates a lie (observable). Assume the following.

1.  If a suspect lies, the lie detector indicates that it is a lie with a probability of 0.8.
2.  If a suspect tells truth, the lie detector indicates that it is a lie with a probability of 0.05.
3.  A suspects lies with a probability of 0.9.

Given the notations $L$ and $D$, the above assumptions can be expressed as

- $P(D \mid L) = 0.8$,

- $P(D \mid L^c) = 0.05$, and

- $P(L) = 0.9$.

If the investigator observes a lie signal, what is the probability that the suspect lied? In other words, what is $P(L \mid D)$?

<!--# Recall the multiplication rule - set it up for both conditional probabilities, set them equal and solve for the unknown probability -->

**Theorem:** For two events $A$ and $B$, the conditional probability of $B$ given $A$ is

$$P(B \mid A) = \frac{ P(A \mid B)\cdot P(B) }{ P(A) }$$

provided that $P(A) > 0$. This statement is known as the **Bayes' Theorem**.

**Recall:** Since $B$ and $B^c$ are a partition of any sample space,

$$P(A) = P(A \mid B) \cdot P(B) + P(A \mid B^c) \cdot P(B^c) \, . $$ Therefore, the Bayes' Theorem can be also written as

$$P(B \mid A) = \frac{ P(A \mid B) \cdot P(B) }{ P(A \mid B) \cdot P(B) + P(A \mid B^c) \cdot P(B^c) } \, .$$ {#eq-bayes}

**Recall**: $P(E^c) = 1 - P(E)$ for any event $E$.

This property holds for a conditional probability. In other words, $P(E^c \mid C) = 1 - P(E \mid C)$ for any possible condition $C$.

Therefore, @eq-bayes can be also written as

$$ P(B \mid A) = \frac{ P(A \mid B) \cdot P(B) }{ P(A \mid B) \cdot P(B) + [1 - P(A^c \mid B^c)] \cdot [1 - P(B)] } \, . $$

***Example*****:** Let us revisit the lie detector scenario. Under the assumptions of $P(D \mid L) = 0.8$, $P(D \mid L^c) = 0.05$, and $P(L) = 0.9$. Apply the Bayes' Theorem (in the form of @eq-bayes) to find the probability that the suspect is lying given the lie detector claims it is a lie, or $P(L \mid D)$:

<!-- $$ P(L \mid D) = \frac{ (0.8)(0.9) }{ (0.8)(0.9) + (0.05)(1-0.9) } = 0.993 \, . $$ -->

In words, if the lie detector indicated that the suspect lied, there is a ________ probability that the suspect really lied.

### Sensitivity and Specificity

Bayes Theorem is commonly applied to scenarios where we only know the probability of a detection under a particular condition, but we need to know the probability fo the condition given a positive detection.  

- In the context of the lie detector scenario, the conditional probability $P(D \mid L)$ is called the **sensitivity**. A high sensitivity implies a high chance of detecting a lie.  

- The conditional probability $P(D^c \mid L^c)$ is called the **specificity**. A high specificity implies a high chance of detecting a truth.  

- The probability $P(L)$ is called the **prevalence**. A high prevalence implies that subjects lie frequently. 

- The conditional probability $P(L \mid D)$ is called the **positive predictive value**. This is the value that tells us how good the lie detector really is!  


@eq-bayes informs that the reliability of a lie detector depends on the sensitivity and specificity (how good the lie detector is), but it also depends on the prevalence (how untruthful or truthful subjects are).  
 
Given the sensitivity ($\texttt{sens}$) and specificity ($\texttt{spec}$), the positive predictive value ($\texttt{ppv}$) is a function of the prevalence ($\texttt{ppv}$) as follows:

$$\texttt{ppv} = \frac{ \texttt{sens} \cdot \texttt{prev} }{ \texttt{sens} \cdot \texttt{prev} + (1 - \texttt{spec}) \cdot (1 - \texttt{prev}) } \, .$$

The below R code demonstrates the relationship between $\texttt{ppv}$ and $\texttt{prev}$ at $\texttt{sens} = 0.8$ and $\texttt{spec} = 0.95$. The graph demonstrates that the positive predictive value $P(B \mid A)$ increases with respect to the prevalence $P(B)$. In other words, we need frequent lies (i.e., $P(B)$ is high) in order for the lie detector to predict lies well (i.e., $P(L \mid D)$ is high).

```{webr}
#| label: pos_pred_val
### below is a created function, called pos_pred_val()
### input: sens, spec, prev
### output: ppv

pos_pred_val <- function( sens, spec, prev ) {
   num <- sens * prev
   den <- sens * prev + ( 1 - spec ) * ( 1 - prev )
   ppv <- num / den
   ppv }

pos_pred_val( sens = 0.8, spec = 0.95, prev = 0.1 ) ### returns ppv for given values
pos_pred_val( sens = 0.8, spec = 0.95, prev = 0.001 ) ### returns ppv for given values
```


Notice how the positive predictive values ($\texttt{ppv}$) changes based on the prevalence of lies.  If the suspect is truthful (low prevalence) how does this affect the ability of the machine to detect a lie?  

  
    
     

Let's observe the positive predictive values across all possible prevalences.  

```{webr}
#| label: ppv-graph
#| fig-height: 5
#| fig-width: 5 
#| fig-alt: "The plot of positive predictive value as a function of prevalence."

est_prev <- seq( 0, 1, 0.001 ) ### creates vector of prevalences from 0 to 1 with an increment of 0.001
est_ppv <- pos_pred_val( sens = 0.8, spec = 0.95, prev = est_prev ) ### calculates ppv across all points

## plots the estimated ppv vs the prevalence as a curve
gf_line( est_ppv ~ est_prev, 
         title = "PPV vs. Prevalence",
         xlab = "Prevalence of Lies", 
         ylab = "Positive Predictive Values for Detection of a Lie" )
```

What do you notice about the performance of the lie detector?



**Example:** Let us continue from the previous example. If the lie detect is quiet (no indication of a lie), what is the probability that the suspect really did not lie? In other words, what is the conditional probability of $L^c$ given $D^c$? Applying the Bayes' Theorem.  

<!-- $$\begin{aligned} -->
<!-- P(L^c \mid D^c) & = \frac{P(D^c \mid L^c) \cdot P(L^c)}{P(D^c)} \\ -->
<!-- & = \frac{P(D^c \mid L^c) \cdot P(L^c)}{P(D^c \mid L^c) \cdot P(L^c) + P(D^c \mid L) \cdot P(L)} \\ -->
<!-- & = \frac{(0.95)(0.1)}{(0.95)(1 - 0.9) + (1-0.8)(0.9)} \\ -->
<!-- & = 0.345 \, . -->
<!-- \end{aligned}$$ -->

Such a low probability implies that we cannot trust the suspect when there is no indication of a lie by the lie detector. It is only useful when the lie detector indicates that it is a lie.



### Prenatal Testing for Rare Diseases

::: callout-note
## Context

These are original, hypothetical exercises inspired by the article's discussion of prenatal genetic screening and false positives for rare conditions. A screening result estimates risk; it does not diagnose a condition. The numbers below are chosen for practice, not for medical decision-making.
:::

#### Bayes' Theorem and Screening

Let (C) denote “the condition is present” and (+) denote “the screening test is positive.” Then

$$P(C \mid +) =
\frac{P(+ \mid C)P(C)}
{P(+ \mid C)P(C) + P(+ \mid \neg C)P(\neg C)}.$$

#### Rare-condition Screening{#rare-condition}

A condition occurs in 1 out of every 2,000 pregnancies. A screening test detects the condition 95% of the time when it is present and has a 1% false-positive rate.

If a patient receives a positive result, what is the probability that the fetus actually has the condition?

::: {.callout-tip collapse="true"}
#### Rare-condition Screening Solution

$$P(C \mid +) = \frac{(.95)(1/2000)}{(.95)(1/2000) + (.01)(1999/2000)} \approx 0.0454.$$
:::

#### Natural-frequency in the Population

Use the information from @rare-condition, but imagine 200,000 pregnancies.

- How many fetuses are expected to have the condition?  
- How many true positives will the test produce?  
- How many false positives will it produce?  
- Of all positive results, what percentage are true positives?  

::: {.callout-tip collapse="true"}
#### Natural-frequency in the Population Solutions

- Of $200,000$ pregnancies, $100$ are expected to have the condition.  

- The test identifies $(0.95(100) = 95)$ true positives.

- Of the 199,900 pregnancies without the condition, $1%$ produce false positives: $(0.01(199,900) = 1,999)$.

- Thus, there are $(95 + 1,999 = 2,094)$ positive results, of which $\frac{95}{2,094} \approx 4.54% $ are true positives.
:::

#### Comparing Two Conditions

Test A screens for Condition A, which occurs in 1 in 100 pregnancies. Test B screens for Condition B, which occurs in 1 in 2,000 pregnancies. Both tests have 99% sensitivity and a 1% false-positive rate.

For each test, calculate $P(\text{condition} \mid \text{positive test})$. Why can equally accurate tests lead to very different interpretations of a positive result?

::: {.callout-tip collapse="true"}
#### Comparing Two Conditions

For Condition A:

$$\frac{(.99)(.01)}{(.99)(.01) + (.01)(.99)} = 0.50.$$

So a positive test gives a **50%** probability that Condition A is present.

For Condition B:

$$\frac{(.99)(.0005)}{(.99)(.0005) + (.01)(.9995)} \approx 0.0472.$$

So a positive test gives only about a **4.72%** probability that Condition B is present. The key difference is the prevalence: Condition B is much rarer.

:::

#### Different Risk Group

In a higher-risk group, a condition occurs in 1 in 250 pregnancies. The same screening test from @rare-condition is used: 95% sensitivity and a 1% false-positive rate.

What is the probability of the condition after a positive result? Compare it with your answer to @rare-condition. What changed, and what stayed the same?

::: {.callout-tip collapse="true"}
#### Different Risk Group Solution

$$\frac{(.95)(.004)}{(.95)(.004) + (.01)(.996)} \approx 0.276.$$

The probability after a positive result is about **27.6%**. The test's sensitivity and false-positive rate stayed the same; the prior probability (prevalence) increased.
:::

#### A Negative Result

A condition occurs in 1 in 500 pregnancies. A screening test has 98% sensitivity and a 2% false-positive rate.

If the result is negative, what is the probability that the fetus still has the condition?

::: {.callout-tip collapse="true"}
#### A Negative Result Solution

The false-negative rate is $$(1 - .98 = .02)$$. Therefore,

$$P(C \mid -) = \frac{(.02)(.002)}{(.02)(.002) + (.98)(.998)} \approx 0.0000409.$$

The probability is approximately **0.0041%**, or about 1 in 24,450.
:::

#### An Advertising Claim

A company advertises that its screening test is “99% accurate.” Suppose the condition occurs in 1 in 10,000 pregnancies, sensitivity is 99%, and specificity is 99%.

- Find the positive predictive value.  
- Is “99% accurate” enough information to tell a patient the chance that a positive result is correct? Explain using Bayes' theorem.  

::: {.callout-tip collapse="true"}
#### An Advertising Claim Solution

$$P(C \mid +) = \frac{(.99)(.0001)}{(.99)(.0001) + (.01)(.9999)} \approx 0.0098.$$

The positive predictive value is about **0.98%**.  

No: a general accuracy claim omits the condition's prevalence, as well as the separate sensitivity and specificity values that Bayes' theorem requires.
:::

#### A Follow-up Diagnostic Test

A screening test yields a 4% probability that a fetus has a condition after a positive result. A follow-up diagnostic test has 99.9% sensitivity and 99.9% specificity.

Assuming the diagnostic test is positive, use Bayes' theorem to estimate the updated probability that the fetus has the condition. Explain why follow-up testing matters.

::: {.callout-tip collapse="true"}
#### A Follow-up Diagnostic Test Solution

Treat the 4% result as the new prior probability, or $P(C)$:

$$P(C \mid +) = \frac{(.999)(.04)}{(.999)(.04) + (.001)(.96)} \approx 0.9765.$$

The updated probability is approximately **97.65%**. A highly accurate follow-up test can substantially reduce uncertainty after a screening result.
:::

#### Source context

The problems are based on the general distinction between prenatal screening and diagnostic testing discussed in [the referenced New York Times article](https://www.nytimes.com/2022/01/01/upshot/pregnancy-birth-genetic-testing.html). For current clinical guidance, consult a qualified health-care professional.

## Bayes Theorem and Machine Learning

Bayes Theorem not only spurred a whole new philosophical approach towards statistical inference (called Bayesian Inference), in contrast to the Frequentist Inference approach we will focus on in our course, but it has also provided the foundation for many machine learning methods.  

### Example 1: A Naive Bayes Spam Classifier

A Naive Bayes classifier uses Bayes' theorem to estimate the probability that a message belongs to a class, such as *spam* or *not spam*, after observing its words. The word *naive* refers to an assumption: once the class is known, the model treats the observed word features as independent.

Suppose that, in a training set:

- 20% of emails are spam: $(P(S) = 0.20)$;
- 75% of spam emails contain the word “winner”: $(P(W \mid S) = 0.75)$;
- 5% of non-spam emails contain “winner”: $(P(W \mid \neg S) = 0.05)$.

For an email that contains “winner,” the classifier calculates

$$P(S \mid W) = \frac{P(W \mid S)P(S)} {P(W \mid S)P(S) + P(W \mid \neg S)P(\neg S)}.$$

$$P(S \mid W) = \frac{(.75)(.20)}{(.75)(.20) + (.05)(.80)} = \frac{.15}{.19} \approx 0.789.$$

So the model estimates a **78.9% probability** that the email is spam. In practice, it combines evidence from many words:

$$P(S \mid w_1, \ldots, w_n) \propto P(S)\prod_{i=1}^{n}P(w_i \mid S).$$

The class with the larger posterior probability is the prediction. This is the same update as in the screening problems: begin with a prior probability, then revise it using evidence.

### Example 2: Bayesian A/B testing

A/B testing is used to identify the best versions/features to include on apps, websites, and other user interfaces. In a Bayesian A/B test, we have version A and version B, and we treat each version's conversion rate (getting the user to click on something) as an unknown probability and update our beliefs as observations arrive. Unlike a classical A/B test, the central result can be stated directly as a probability about the two versions.

Suppose a website runs two versions of a sign-up page. Before collecting data, use a uniform prior for each conversion rate, that is both versions have the same probability of working. 

After the experiment:

| Version | Visitors | Sign-ups | Non-sign-ups |
|---------|---------:|---------:|-------------:|
| A       |      100 |       12 |           88 |
| B       |      100 |       18 |           82 |

We then take that data and update our assumption and find that 


The posterior means, $E$ or expected value, are

$$E[p_A \mid \text{data}] = \frac{13}{102} \approx 12.7%$$ and
$$E[p_B \mid \text{data}] = \frac{19}{102} \approx 18.6%. $$

To make a decision, draw many samples from the two posterior distributions and count how often $p_B > p_A$. That fraction estimates $P(p_B > p_A \mid \text{data})$: the probability, given the model and observed data, that Version B has the higher conversion rate. This illustrates Bayes' theorem as a learning process—prior beliefs are updated to posterior beliefs as data accumulate.

## Wrap-Up

In your groups work on the following problems. You can check your answers below. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Be ready to show your work on the board.

#### Problem 1
Assume $P(A \mid B) = 0.9$, $P(A \mid B^c) = 0.3$, and $P(B) = 0.7$. Calculate the following probabilities.

a. $P(B^c)$  

```{r}
#| label: wrap-1-a
#| echo: false
checkdown::check_question(answer = 0.3,
          button_label = "Check Your Answer")
```

b. $P(A)$  

```{r}
#| label: wrap-1-b
#| echo: false
checkdown::check_question(answer = 0.72,
          button_label = "Check Your Answer")
```

c. $P(A^c)$  

```{r}
#| label: wrap-1-c
#| echo: false
checkdown::check_question(answer = 0.28,
          button_label = "Check Your Answer")
```

d. $P(A^c \mid B)$  

```{r}
#| label: wrap-1-d
#| echo: false
checkdown::check_question(answer = 0.1,
          button_label = "Check Your Answer")
```

e. $P(A^c \mid B^c)$  

```{r}
#| label: wrap-1-e
#| echo: false
checkdown::check_question(answer = 0.7,
          button_label = "Check Your Answer")
```

f. $P(B \mid A)$  

```{r}
#| label: wrap-1-f
#| echo: false
checkdown::check_question(answer = 0.875,
          button_label = "Check Your Answer")
```

g. $P(B^c \mid A)$  

```{r}
#| label: wrap-1-g
#| echo: false
checkdown::check_question(answer = 0.125,
          button_label = "Check Your Answer")
```

h. $P(B^c \mid A^c)$  

```{r}
#| label: wrap-1-h
#| echo: false
checkdown::check_question(answer = 0.75,
          button_label = "Check Your Answer")
```

i. $P(B \mid A^c)$  

```{r}
#| label: wrap-1-i
#| echo: false
checkdown::check_question(answer = 0.25,
          button_label = "Check Your Answer")
```

#### Problem 2
Suppose any news is good or bad. If there is good news in the morning, there is good news in the evening with a probability of 0.75. If there is bad news in the morning, there is bad news in the evening with a probability of 0.25. Assume morning news is good 51% of the time. Suppose we missed morning news today, and there is good news in the evening. What is the probability that today's morning was good?

```{r}
#| label: wrap-2
#| echo: false
checkdown::check_question(answer = 0.51,
          button_label = "Check Your Answer")
```

#### Problem 3
Suppose we toss a coin inside a room. If it rains outside, head occurs with a probability of 0.99. If it does not rain outside, head occurs with a probability of 0.1. Suppose it rains 10% of the time. Given head, calculate the probability that it is raining outside.

```{r}
#| label: wrap-3
#| echo: false
checkdown::check_question(answer = 0.5238,
          button_label = "Check Your Answer")
```

#### Problem 4
Suppose we toss a coin inside a room. If it rains outside, head occurs with a probability of 0.99. If it does not rain outside, head occurs with a probability of 0.1. Suppose it rains 10% of the time. Given head, calculate the probability that it is raining outside.

```{r}
#| label: wrap-4
#| echo: false
checkdown::check_question(answer = 0.0909,
          button_label = "Check Your Answer")
```

## Assignment
Check your answer by inputting them below their respective question. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified.

Though you are checking the numeric answers here, you should make sure your written work is complete and your mathematical notation is correct as both will be graded on the homework quiz.

#### Problem 1
Assume $P(A) = 0.5$, $P(B \mid A) = 0.8$, and $P(B^c \mid A^c) = 0.7$. Calculate the following probabilities.

a. $P(A^c)$  

```{r}
#| label: asn-1-a
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

b. $P(B)$  

```{r}
#| label: asn-1-b
#| echo: false
checkdown::check_question(answer = 0.55,
          button_label = "Check Your Answer")
```

c. $P(B^c)$  

```{r}
#| label: asn-1-c
#| echo: false
checkdown::check_question(answer = 0.45,
          button_label = "Check Your Answer")
```

d. $P(B \mid A^c)$  

```{r}
#| label: asn-1-d
#| echo: false
checkdown::check_question(answer = 0.3,
          button_label = "Check Your Answer")
```

e. $P(B^c \mid A^c)$  

```{r}
#| label: asn-1-e
#| echo: false
checkdown::check_question(answer = 0.7,
          button_label = "Check Your Answer")
```

f. $P(A \mid B)$  

```{r}
#| label: asn-1-f
#| echo: false
checkdown::check_question(answer = 0.7273,
          button_label = "Check Your Answer")
```

g. $P(A^c \mid B)$  

```{r}
#| label: asn-1-g
#| echo: false
checkdown::check_question(answer = 0.2727,
          button_label = "Check Your Answer")
```


h. $P(A \mid B^c)$  

```{r}
#| label: asn-1-h
#| echo: false
checkdown::check_question(answer = 0.2222,
          button_label = "Check Your Answer")
```

i. $P(A^c \mid B^c)$  

```{r}
#| label: asn-1-i
#| echo: false
checkdown::check_question(answer = 0.7778,
          button_label = "Check Your Answer")
```

j. $P(A \mid A^c)$  

```{r}
#| label: asn-1-j
#| echo: false
checkdown::check_question(answer = 0,
          button_label = "Check Your Answer")
```

k. $P(A \mid A)$  

```{r}
#| label: asn-1-k
#| echo: false
checkdown::check_question(answer = 1,
          button_label = "Check Your Answer")
```

#### Problem 2
In an airport, a body scanner beeps when it detects a metal (a potential concealed weapon). If a traveler enters the body scanner with a metal, it beeps with a probability 0.999. If a traveler enters the body scanner without a metal, it beeps (false alarm) with a probability of 0.1.

a. Assume 0.01% of travelers (1 in 10,000) enter the body scanner with a metal. If a body scanner beeps, what is the probability that the traveler actually has a metal?(Round to **6 decimal places** for your response.)  

```{r}
#| label: asn-2-a
#| echo: false
checkdown::check_question(answer = 0.000998,
          button_label = "Check Your Answer")
```

b. Repeat part a assuming 10% of travelers enter the body scanner with a metal.  

```{r}
#| label: asn-2-b
#| echo: false
checkdown::check_question(answer = 0.526,
          button_label = "Check Your Answer")
```

#### Problem 3 
For fair criminal trials, defendants are assumed to be innocent until proven guilty, and prosecutors must provide sufficient evidence in order to conclude that defendants are guilty. Assume the following:

- If a defendant is innocent, it is concluded that the defendant is guilty with a probability of 0.05.  
- If a defendant is guilty, it is concluded that the defendant is guilty with a probability of 0.9.  

In a hypothetical scenario, suppose prosecutors accuse any people so that 99% of defendants are innocent. If the conclusion is guilty, what is the probability that the defendant is truly guilty?

```{r}
#| label: asn-3
#| echo: false
checkdown::check_question(answer = 0.1538,
          button_label = "Check Your Answer")
```

#### Problem 4 
Until 2013, all employees in Canadian workplaces were subject to drug test by a random selection.^1^ The drug test was intended to identify subjects who use illegal drugs, but it had been criticized that too many innocent employees were tested. In 2013, the Supreme Court of Canada ruled that there must be a reasonable cause (e.g., suspicious actions) to drug test an employee. This new policy would narrow the target population of drug test from all employees to suspicious employees. Let $U$ denote a drug use in a target population and $T$ denote a postive drug test result. Assume both sensitivity and specificity of the drug test have been constant at 0.95, and assume that only 5% are drug users among all employees and 95% are drug users among those who act suspiciously. Calculate the following.

a. $P(T \mid U)$  

```{r}
#| label: asn-4-a
#| echo: false
checkdown::check_question(answer = 0.95,
          button_label = "Check Your Answer")
```

b. $P(T \mid U^c)$  

```{r}
#| label: asn-4-ba
#| echo: false
checkdown::check_question(answer = 0.05,
          button_label = "Check Your Answer")
```

c. $P(T^c \mid U^c)$  

```{r}
#| label: asn-4-c
#| echo: false
checkdown::check_question(answer = 0.95,
          button_label = "Check Your Answer")
```

d. $P(T^c \mid U)$  

```{r}
#| label: asn-4-d
#| echo: false
checkdown::check_question(answer = 0.05,
          button_label = "Check Your Answer")
```

e. $P(U)$ under the old policy (prior to 2013)  

```{r}
#| label: asn-4-e
#| echo: false
checkdown::check_question(answer = 0.05,
          button_label = "Check Your Answer")
```

f. $P(U)$ under the new policy  

```{r}
#| label: asn-4-f
#| echo: false
checkdown::check_question(answer = 0.95,
          button_label = "Check Your Answer")
```

g. Under the [old]{.underline} policy, what is the probability that the subject is [not a drug user]{.underline} when the test result is [positive]{.underline}? (i.e., accuse an innocent subject)  

```{r}
#| label: asn-4-g
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

h. Under the [old]{.underline} policy, what is the probability that the subject is [a drug user]{.underline} when the test result is [negative]{.underline}? (i.e., free a guilty subject)  

```{r}
#| label: asn-4-h
#| echo: false
checkdown::check_question(answer = 0.0028,
          button_label = "Check Your Answer")
```

i. Under the [new]{.underline} policy, what is the probability that the subject is [not a drug user]{.underline} when the test result is [positive]{.underline}?  

```{r}
#| label: asn-4-i
#| echo: false
checkdown::check_question(answer = 0.0028,
          button_label = "Check Your Answer")
```

j. Under the [new]{.underline} policy, what is the probability that the subject is [a drug user]{.underline} when the test result is [negative]{.underline}?  

```{r}
#| label: asn-4-j
#| echo: false
checkdown::check_question(answer = 0.5,
          button_label = "Check Your Answer")
```

k. Multiple-choice: All employees had been targeted under the old policy, and only suspicious employees have been targeted under the new policy. If the test result is negative, the subject is concluded to be clean, and will not be tested for the next three years. If the test result is positive, the subject is temporarily concluded to be a potential risk in the workplace, and will be observed until proven to be clean. According to the probability calculations in previous parts, the \[...\] policy raises the concern of safety in the workplace, and the \[...\] policy raises the concern of privacy.  


```{r}
#| label: asn-4-k
#| echo: false
checkdown::check_question(answer = "new; old",
                          options = c("old; old",
                                      "old; new",
                                      "new; old",
                                      "new; new",
                                      "both; both"),
                          button_label = "Check Your Answer", 
                          type = "radio")
```


^1^ Hurley, W. J. (2015). Mandatory drug testing in the Canadian workplace: A note on the recent supreme court decision and its effect on the misclassification rate of non-drug users, Chance, 28(1), 21-26. \url{http://chance.amstat.org/2015/02/drug-testing/}


#### Problem 5 
Imagine you are the wife of a man who was in Titanic which sank in the morning of April 15, 1912. News on radio said 30% of all passengers survived, 25% were men among survivors, and 80% were men among non-survivors. You do not know whether your husband survived or not.

a. What percent of passengers were men?  

```{r}
#| label: asn-5-a
#| echo: false
checkdown::check_question(answer = 0.635,
          button_label = "Check Your Answer")
```


b. What is the probability that your husband survived?  

```{r}
#| label: asn-5-b
#| echo: false
checkdown::check_question(answer = 0.1181,
          button_label = "Check Your Answer")
```

