---
title: "Summarizing and Visualizing Data"
format: 
  live-html:
    webr:
      packages: 
        - ggformula
        - mosaic
        - tidyverse
      repos:
        - https://projectmosaic.r-universe.dev
      resources:
        - https://raw.githubusercontent.com/jcanner/math-320/refs/heads/master/note-outlines/8-intro-to-stats/study-hours.csv
        - https://raw.githubusercontent.com/jcanner/math-320/refs/heads/master/note-outlines/8-intro-to-stats/king.csv
      cell-options:
        warning: true
        completion: false
execute: 
  warning: false 
  message: false
  echo: true
embed-resources: false
code-copy: true
code-overflow: wrap
toc: true
number-sections: true
css: ../styles.css
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}\
{{< include ../_extensions/r-wasm/live/_gradethis.qmd >}}

```{webr}
#| label: theme
#| include: false
#| edit: false
library(ggplot2)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

```{r}
#| label: setup-checkdown
#| include: false
library(checkdown)
library(ggformula)
library(mosaic)
library(tidyverse)
theme_set(theme_light())
options(ggplot2.continuous.color = scale_color_viridis_c, 
        ggplot2.continuous.fill = scale_fill_viridis_c,
        ggplot2.discrete.color = scale_color_viridis_d,
        ggplot2.discrete.fill = scale_fill_viridis_d)
```

## Good Practice of Statistics

A good practice of statistics is as follows.

- Set clear objective(s).

- Collect data which can address the objective(s).

- Calculate informative, concise, and interpretable descriptive statistic(s).

- Present the results in table(s) and graphic(s) which are self-explanatory.

- Interpret the results in the context.

## Categorical Data

Categorical data can be summarized by the **count** of each level, and then we can calculate the **percent** or **proportion** of each level as needed. The categorical data can be visualized by a **barplot**.


We will use the `study_hours` data again, so we will read it in and add in our `mutate()` function to set our levels for `year` and our new variable `hours_units`.  
```{r}
#| label: read-study-hours
study_hours <- read_csv("8-intro-to-stats/study-hours.csv", 
                       show_col_types = FALSE) |> 
  mutate(year = factor(year, levels = c("Sophomore",
                                        "Junior",
                                        "Senior")),
         hours_units = hours/units)
```

### Counts, Proportions, and Percentages

The `tally()` function allows us to summarize categorical data via counts, proportions, or percentages, both as single variable and looking across multiple variables, if you specify the format as "percent", "proportion", or "count". The default is "count".

The `formula` can be one of the following -

- `~x` for a single variable x
- `y ~ x` for two variables, `y` will be conditioned on `x` for non-count tables.  
- `~x | y` for a conditional calculation
- 

```{r}
#| label: stats-cat

#percentage, formula ~x
tally(~ year, data = study_hours, format = "percent")

#proportions - can use "proportion" or "prop"
tally(~ gender, data = study_hours, format = "prop")

#count table with two variables, formula y~x
tally(gender ~ year, data = study_hours, format = "count")
```


### Bar plots

We can make a bar plot for a single variable using the following functions:  

- `gf_bar(~x, data = mydata)` for unsummarized counts of `x`  
- `gf_col(y ~ x, data = mydata)` for summarized counts `y` of `x`  
- `gf_props(~x, data = mydata)` for unsummarized proportions of `x`  
- `gf_percents(~x, data = mydata)` for unsummarized percentages of `x`  

```{r}
#| label: single-barplots
gf_bar(~gender, data = study_hours,
       xlab = "Gender",
       ylab = "Number of Students",
       title = "Sample of n = 110 University Students in the College of Science")

gf_props(~year, data = study_hours,
       xlab = "Class Rank",
       ylab = "Proportion of Students",
       title = "Sample of n = 110 University Students in the College of Science")
```

Always be sure to provide labels for each axis (`xlab`, `ylab`) and an informative `title` for each graph that includes: 

- Variable Names and Units, if appropriate
- Summary measures displayed on the graph
- Information about the sample or population, including size



### Conditional Proportions

A contingency table is a simple way of looking at the relationship between two categorical variables.


```{r}
#| label: contingency-table
tally(gender ~ year, data = study_hours)
```

A conditional proportion is the proportion calculated over a specific subgroup - it is the empirical analog to a conditional probability $P(A | B)$.  The formula can take two forms: 

- `y ~ x`, which is `y` conditioned on `x` or
- `~ y | x`, which is also `y` conditioned on `x`

I prefer the `~ y | x` format because it is the same format that will be needed for an analogous barplot of the conditional outcomes.  


```{r}
#| label: conditional-proportion
tally(~ gender | year, data = study_hours, format = "prop")
```

What does the table tell us about the distribution of female-identifying students across the class years in our data?  






We can visualize our conditional proportions using the `gf_props()` function.  Notice the new argument `fill = ~gender` which will fill our bars with a color for each level in `gender`.  

```{r}
#| label: conditional-bar-plot
gf_props(~gender | year, data = study_hours, 
         fill = ~gender, #fills bars with a color based on gender
         xlab = "Gender",
         ylab = "Proportion of Students",
         title = "Sample of n = 110 University Students in the College of Science",
         show.legend = FALSE) #hides unnecessary legend 
```




### Case Study I: Titanic

Titanic was a British passenger liner. It sank in the North Atlantic Ocean in the morning of April 15, 1912. The cause was collision with an iceberg, and many passengers lost lives. `titanic.csv` is a dataset which includes information about 1,316 passengers. There are five columns:

- `id`: passenger identification number

- `class`: passenger class (first, second, or third)

- `age`: passenger’s age group (adult or child)

- `sex`: passenger’s sex (female or male)

- `survived`: passenger’s survival status (no or yes)

The objective is to investigate if certain groups of passengers had higher chances of survival.

[Cal ICOR Link](https://csumb.jupyter.cal-icor.org/hub/user-redirect/git-pull?repo=https%3A%2F%2Fgithub.com%2Fjcanner%2Fmath-320-activities&branch=main&urlpath=rstudio%2F)

[R On Computer: Download Zip File](9-summarizing-data/summarizing-data-activity.zip)

Our goal is to produce statistics and graphics that help us explore the data on survival, similar to the output below, using R. 

|     Variable    |   Level  | Did not survive | Survived  | Total |
|:---------------:|:--------:|:---------------:|:---------:|:-----:|
|       Sex       |  Female  |    123 (28%)    | 324 (72%) |  447  |
|                 |   Male   |    694 (80%)    | 175 (20%) |  869  |
|       Age       |  Adults  |    765 (63%)    | 442 (37%) | 1,207 |
|                 | Children |    52 (48%)     | 57 (52%)  |  109  |
| Passenger Class |  First   |    122 (38%)    | 203 (62%) |  325  |
|                 |  Second  |    167 (59%)    | 118 (41%) |  285  |
|                 |  Third   |    528 (75%)    | 178 (25%) |  706  |
|      Total      |  Total   |    817 (62%)    | 499 (38%) | 1,316 |

: Count and percent of passengers who survived and who did not survive by sex, age, and passenger class








## Numeric Data

Numeric data can be summarized by sample mean and standard deviation as well as 5-number summary. The 5-number summary can be used to draw a boxplot which is a quick visual summary of numeric data. To do that, we first must define **outliers**.

### Outliers

Let $X_{(1)}$, $Q_1$, $M$, $Q_3$, and $X_{(n)}$ be the sample minimum, first quartile, median, third quartile, and maximum, respectively. The **inter-quartile range** (or **IQR**) is defined as

$$
\texttt{IQR} = Q_3 - Q_1 \, .
$$

The **lower fence** and **upper fence** are defined as

$$
\begin{aligned}
\texttt{LF} & = Q_1 - 1.5 \times \texttt{IQR} \, , \\
\texttt{UF} & = Q_3 + 1.5 \times \texttt{IQR} \, ,
\end{aligned}
$$

any data values outside of $(\texttt{LF}, \texttt{UF})$ are **outliers**. (In general, any extremely low or high data values are referred to as outliers, and it is an arbitrary definition of outliers.) 



### Boxplot

Once we know the $(\texttt{LF}, \texttt{UF})$ we can use this to draw the boxplot by defining our "whiskers".  

The **lower whisker** and **upper whisker** are defined as follows:

- Among data values greater than or equal to `LF`, let `LW` be the closest data value to `LF`. Then `LW` is called the **lower whisker**.

- Among data values less than or equal to `UF`, let `UW` be the closest data value to `UF`. Then `UW` is called the upper whisker.

- The outliers will then be represented as dots beyond the last value on the whisker.

- The "box" is then made up of $Q_1$, $M$, and $Q_3$ values. 


R will do these calculations for us within the `gf_boxplot()` function:

```{r}
#| label: boxplot

gf_boxplot(~hours_units, data = study_hours,
           xlab = "Study Hours Per Week Per Unit",
           title = "Study Habits of Sudents in the College of Science (n = 110) at the University") |> 
  gf_theme(axis.ticks.y = element_blank(),  #removes y-axis ticks
           axis.text.y = element_blank())   #removes y-axis labels

```

We can also compare side-by-side boxplots horizontally by specifying the categorical variable as the `y` position and the numeric variable as the `y` position:

```{r}
#| label: boxplot-horizontal
gf_boxplot(gender ~ hours_units, data = study_hours,
           xlab = "Study Hours Per Week Per Unit",
           ylab = "Gender",
           title = "Study Habits of Sudents in the College of Science (n = 110) at the University") 
```
We can also flip it vertically by switching our categorical variable to the `x` position rather than `y` position. 
```{r}
#| label: boxplot-vertical
gf_boxplot(hours_units ~ year, data = study_hours,
           ylab = "Study Hours Per Week Per Unit",
           xlab = "Class Year",
           title = "Study Habits of Sudents in the College of Science (n = 110) at the University") 
```


#### Descriptions of Shape - Skew

In looking at a boxplot we can also evaluate the "skew" of the distribution as one of the following:

- Right/Positive Skew




- Left/Negative Skew   




- Symmetric




Let's compare the summary statistics across gender for hours per week per unit.  The function `df_stats()` gives us the five number summary plus mean/sd.    


```{r}
#| label: sum-stats-skew
df_stats(hours_units ~ gender, data = study_hours )

```

How do you think skew impacts our measures of center? 






### Histogram

A histogram is another way of visualizing numeric data. It partitions the range of observed data by creating a bin $B_i = [b_{i-1}, b_i)$ for $i = 1, 2, \dots, m$; counts data values belonging to each bin; and uses a graphic like a barplot to visualize the relative frequency of each bin. Below is a histogram of study hours per week per unit with a bin width of 0.25.


```{r}
#| label: histogram

gf_histogram(~hours_units, data = study_hours,
             xlab = "Study Hours Per Week Per Unit",
             binwidth = 0.25,
             color = "black") #color outlines the bars

```


We can adjust the bin width to whatever is reasonable to create a clear visualization - it is balance between too much "granularity" and too little.  Here is a bin width of 0.20 - which histogram do you think is better?

```{r}
#| label: histogram-binwidth-change

gf_histogram(~hours_units, data = study_hours,
             xlab = "Study Hours Per Week Per Unit",
             binwidth = 0.20,
             color = "black") #color outlines the bars

```


#### Descriptions of Shape - Modality

Just like the boxplot, we can also describe the skew of the data distribution based on the histogram (and id outliers), but histograms also provide information on another characteristics of shape - modality, or how many peaks a distribution might have.

Draw a unimodal, bimodal, and multimodal with different skews:












### Case Study II: European Soccer {#sec-euro}

Have European soccer players become faster over the 8 years (from Euro 2016 to 2024)? Of course, soccer players do not generate their maximum speed every minute, but they run at their maximum speed at critical moments. Exercise scientists believe that the maximum speed (km/h) is an indicator of athletes' physical strength and speed. Our goal will be to reproduce the statistics below as well as create visualizations for comparisons to see how maximum speed has changed over time in the Euro Cups.


| Year | Min. |  Q1  | Median |  Q3  | Max. | Mean |  SD  |  N  |
|:----:|:----:|:----:|:------:|:----:|:----:|:----:|:----:|:---:|
| 2016 | 14.1 | 26.3 |  28.2  | 29.6 | 32.8 | 27.7 | 2.77 | 429 |
| 2020 | 0.0  | 27.2 |  29.2  | 30.8 | 33.8 | 28.5 | 3.76 | 490 |
| 2024 | 20.5 | 30.1 |  31.8  | 33.1 | 36.5 | 31.3 | 2.60 | 493 |

: Descriptive statistics of maximum speed (km) by the three events



## Wrap-up

In your groups work on the following problems. You can check your answers below. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified. Be ready to show your work on the board.

#### Problem 1  
Fill in the blank.

a. A categorical variable can be visualized by a \[b...\].

b. A numeric variable can be visualized by a \[b...\] using the 5-number summary.

c. A numeric variable can be visualized by a \[h...\] by specifying the bins (intervals) of possible values.

d. Data values which are outside of lower fence and upper fence called \[o...\].

e. In a boxplot, the middle bar inside the box is the sample \[m...\].

f. In a boxplot, the left side of the box is the sample \[f... q...\].

g. In a boxplot, the right side of the box is the sample \[t... q...\].

h. In a boxplot, the left bar and right bar marked outside of the box are the lower \[w...\] and upper \[w...\]. (They are the same word, so answer only once.)

#### Problem 2  
Given a sample of $(1, 2, \dots, 100)$ of size 100, calculate the following statistics based on the below descriptive statistics.

```{r}
#| label: wrap-2
x <- data.frame(x = 1:100)
head(x)
df_stats(~x, data = x)
```

a. inter-quartile range ($\texttt{IQR}$)

b. lower fence ($\texttt{LF}$)

c. upper fence ($\texttt{UF}$)

d. lower whisker ($\texttt{LW}$)

e. upper whisker ($\texttt{LW}$)

f. the number of outliers

#### Problem 3  
Given a sample of (1, 10, 100, 1000, 10000), calculate the following statistics based on the below descriptive statistics.

```{r}
#| label: wrap-3
x <-  data.frame(x = 10 ^ (0:4))
head(x)
df_stats(~x, data = x)
```

a. inter-quartile range ($\texttt{IQR}$)

b. lower fence ($\texttt{LF}$)

c. upper fence ($\texttt{UF}$)

d. lower whisker ($\texttt{LW}$)

e. upper whisker ($\texttt{LW}$)

f. the number of outliers



## Assignment

Check your answer by inputting them below their respective question. All calculations (numeric answers) must be rounded to four decimals (e.g., 0.6667 instead of 2/3, 0.5 or 0.5000 instead of 1/2) unless specified.

Though you are checking the numeric answers here, you should make sure your written work is complete and your mathematical notation is correct as both will be graded on the homework quiz.

#### Problem 1  
This problem is continued from Case Study II (\@sec-euro). Below is a summary of the maximum speed by event within each position in @tbl-euro2. Fill in the blank by *rounding to one decimal place*. (You must show your code.)

::: #tbl-euro2

|  Position  | Year |  Q1  | Median |  Q3  |  Max  | Mean  | SD  |  N  |
|:----------:|:----:|:----:|:------:|:----:|:-----:|:-----:|:---:|:---:|
|  Defender  | 2016 | 26.8 | \[a\]  | 29.8 | 32.1  | 28.1  | 2.4 | 147 |
|            | 2020 | 28.4 |  29.9  | 31.3 | \[b\] | 29.5  | 2.6 | 165 |
|            | 2024 | 31.0 |  32.2  | 33.5 | 35.7  | \[c\] | 2.1 | 151 |
| Midfielder | 2016 | 26.5 |  28.1  | 29.6 | 31.9  | \[d\] | 2.2 | 164 |
|            | 2020 | 27.2 |  29.0  | 30.4 | \[e\] | 28.3  | 4.0 | 193 |
|            | 2024 | 30.1 | \[f\]  | 32.8 | 35.8  | 31.2  | 2.2 | 182 |
|  Forward   | 2016 | 26.2 | \[g\]  | 29.8 | 32.8  | 27.9  | 2.4 | 94  |
|            | 2020 | 27.4 |  29.5  | 31.3 | \[h\] | 29.1  | 3.1 | 106 |
|            | 2024 | 30.5 |  32.1  | 33.2 | 36.5  | \[i\] | 2.3 | 131 |

: Descriptive statistics of maximum speed (km) by event within each position

:::

#### Problem 2  
`affairs.csv` is a partial dataset of a magazine survey conducted in 1969 by Psychology Today. Readers of the magazine were asked to mail in their answers. About 20,000 replies were received, and a subset of them were recorded as a sample. It contains the following variables. In this problem, the goal of data analysis is to study factors associated with the likelihood of extramarital affairs.

- `affairs`: Did you have extramarital affairs during the past year? (`no` or `yes`)

- `gender`: Are you a female or a male? (`female` or `male`)

- `children`: Do you have children? (`no` or `yes`)

- `religiousness`: How religious are you? (`anti`, `not at all`, `slightly`, `somewhat`, or `very`)

- `rating`: How happy are you with your marriage? (`very unhappy`, `unhappy`, `average`, `above average`, or `very happy`)

Reference: Fair, R.C. (1978). A Theory of Extramarital Affairs. Journal of Political Economy, 86, 45–61. @tbl-affair is a summary of the percent of respondents who had extramarital affairs during the past year. Fill in the blank by *rounding to one decimal place*. (You must show your code.)

|   |   | No affairs (%) | At least one affair (%) | N |
|:--:|:--:|:--:|:--:|:--:|
| Gender | Female | 77.1 | 22.9 | 315 |
|  | Male | 72.7 | 27.3 | 286 |
| Children | No | 84.2 | 15.8 | \[a\] |
|  | Yes (1 or more) | 71.4 | 28.6 | \[b\] |
| Religiousness | Anti-religious | \[c\] | \[d\] | 48 |
|  | Not at all | 75.0 | 25.0 | 164 |
|  | Slightly religious | 66.7 | 33.3 | 129 |
|  | Somewhat religious | 82.6 | 17.4 | 190 |
|  | Very religious | \[e\] | \[f\] | 70 |
| Marriage rating | Very unhappy | 50.0 | 50.0 | 16 |
|  | Unhappy | 50.0 | 50.0 | 66 |
|  | Average | 71.0 | 29.0 | 93 |
|  | Above average | 75.3 | 24.7 | 194 |
|  | Very happy | \[g\] | \[h\] | 232 |
| Total | Total | 75.0 | 25.0 | \[i\] |

: Percent of respondents who had extramarital affairs during the past year

#### Problem 3  
Reproduce the following boxplots exactly. (You must show your code.) You can produce each boxplot separately using the `filter()` function or you can use faceting features of graphs.  

```{r}
#| label: asn-3-recreate
#| warning: false
#| echo: false

euro <- read_csv("9-summarizing-data/euro-cup.csv")

gf_boxplot(event ~ top_speed | position, data = euro,
           xlab = "Top Speed (km/hr) of Players",
           ylab = "Euro Cup Year")

```

#### Problem 4  
Reproduce the following barplots exactly. (You must show your code.) For the second barplot, note that `rating` is an ordinal categorical variable. Specify the order of the five levels of `rating` as follows: `very unhappy`, `unhappy`, `average`, `above average`, and `very happy`.

```{r}
#| label: asn-4-data
#| warning: false
#| echo: false
affair <- read_csv("9-summarizing-data/affairs.csv")
tally(~rating, data = affair) ### alphabetical order (default)
affair <- affair |> 
  mutate(rating = factor(rating,
                        levels = c("very unhappy", "unhappy", "average", "above average", "very happy") ))
```

```{r}
#| label: asn-4-plots-1
#| echo: false
#| warning: false

gf_percents(~affairs | gender, data = affair,
            fill = ~affairs,
            xlab = "Did the Individual Have an Affair?",
            ylab = "Had at least one affair (%)",
            title = "Comparison of Gender for those that Extramarital Affairs During the Past Year",
            subtitle = "A total of 315 females and 286 males participated in the survey",
            show.legend = FALSE)

```

```{r}
#| label: asn-4-plots-2
#| echo: false
#| warning: false

gf_percents(~affairs | rating, data = affair,
            fill = ~affairs,
            xlab = "Did the Individual Have an Affair?",
            ylab = "Had at least one affair (%)",
            title = "Likelihood of Extramarital Affairs During the Past Year Given Happiness",
            subtitle = "Included very unhappy (16), unhappy (66), average (93), above average (194) and very happy (232)",
            show.legend = FALSE)

```
